```{r}
#Clear everything out, its probably messy from the data prep file
#rm(list = ls())

#Packages we will need
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(ggplot2)      #data visualization
library(stringr)      #reading information on strings
library(secr)         #the actual model engine
library(rgdal)        #dealing with spatial data

#Set the working directories to read in the files we need
input_directory<-reproducible::checkPath(file.path(getwd(), "2. SECR Prep/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "3. SECR Models/outputs"), create = TRUE)



```


```{r}


```

Here I am going to try and fit a two year multisession model with the 2021 and 2022 data
BUT ALSO READ IN THE HABITAT MASKS!!!
```{r}
#read in camera trap files - order matters here!
camera_names = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

#Prepare the capthist object
deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_reduced.txt"), 
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession, terse = TRUE)


#Ok now try to add the unmarked sightings as you would normally? 

deer_multisession<-addSightings(deer_multisession, unmarked = file.path(input_directory, "unmarked_matrix.txt"), skip = 1)

#Load in the masks, ORDER MATTERS
load(file.path(input_directory, "mask_2021.RData"))
load(file.path(input_directory, "mask_2022.RData"))
oakbay_masks<-list(mask_2021, mask_2022)

start_time<-Sys.time()
#Fit the model
multisession_fit<-secr.fit(deer_multisession, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = oakbay_masks,
                           details = list(autoini = 2), #Was having getting started unless I specify
                           #buffer = 500,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )
#I just like to know how long this takes
end_time<-Sys.time()
end_time - start_time

#Inspect results
summary(multisession_fit)

#Extract estimate (deer/ha) for each session and convert to deer per km2
mean_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["estimate"]][[1]]*100
mean_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["estimate"]][[1]]*100

#extract confidence intervals for each year
upper_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["ucl"]][[1]]*100
lower_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["lcl"]][[1]]*100

upper_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["ucl"]][[1]]*100
lower_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["lcl"]][[1]]*100

#Wrap that into a dataframe and plot
density_estimates<-data.frame(
    year = c(2021, 2022),
    mean_density = c(mean_2021, mean_2022),
    upper_cl = c(upper_2021, upper_2022),
    lower_cl = c(lower_2021, lower_2022)
)

ggplot(density_estimates, aes(x = as.factor(year), y = mean_density))+
  geom_point()+
  geom_errorbar(aes(ymin = lower_cl, ymax = upper_cl), width = 0.2)+
  ylab("Mean Density\u00B1 95% CI")+
  xlab("Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 12))

```



STARTING FRESH

Here I am going to try and fit a two year multisession model with the 2021 and 2022 data
```{r}
#read in camera trap files - order matters here!
camera_names = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

#Prepare the capthist object
deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_reduced.txt"), 
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession, terse = TRUE)


#Ok now try to add the unmarked sightings as you would normally? 

deer_multisession<-addSightings(deer_multisession, unmarked = file.path(input_directory, "unmarked_multisession.txt"), skip = 1)



start_time<-Sys.time()
#Fit the model
multisession_fit<-secr.fit(deer_multisession, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2), #Was having getting started unless I specify
                           buffer = 500,
                           trace = TRUE)
#I just like to know how long this takes
end_time<-Sys.time()
end_time - start_time

#Inspect results
summary(multisession_fit)

#Extract estimate (deer/ha) for each session and convert to deer per km2
mean_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["estimate"]][[1]]*100
mean_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["estimate"]][[1]]*100

#extract confidence intervals for each year
upper_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["ucl"]][[1]]*100
lower_2021<-summary(multisession_fit)[["predicted"]][["session = Oakbay2021"]][["lcl"]][[1]]*100

upper_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["ucl"]][[1]]*100
lower_2022<-summary(multisession_fit)[["predicted"]][["session = Oakbay2022"]][["lcl"]][[1]]*100

#Wrap that into a dataframe and plot
density_estimates<-data.frame(
    year = c(2021, 2022),
    mean_density = c(mean_2021, mean_2022),
    upper_cl = c(upper_2021, upper_2022),
    lower_cl = c(lower_2021, lower_2022)
)

ggplot(density_estimates, aes(x = as.factor(year), y = mean_density))+
  geom_point()+
  geom_errorbar(aes(ymin = lower_cl, ymax = upper_cl), width = 0.2)+
  ylab("Mean Density\u00B1 95% CI")+
  xlab("Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 12))

```

TRYING IT ALL IN ONE!!!

```{r}
#read in camera trap files - order matters here!
camera_names = c(file.path(input_directory, "deer_cameras_2018.txt"),
                 file.path(input_directory, "deer_cameras_2020.txt"),
                 file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

#Prepare the capthist object
deer_multisession_all<-read.capthist(captfile = file.path(input_directory, "deer_multisession.txt"), 
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession_all, terse = TRUE)

#Ok now try to add the unmarked sightings as you would normally? 
deer_multisession_all<-addSightings(deer_multisession_all, unmarked = file.path(input_directory, "unmarked_multisession.txt"), skip = 1)

start_time<-Sys.time()
#Fit the model
multisession_fit_all<-secr.fit(deer_multisession_all, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2), #Was having getting started unless I specify
                           buffer = 500,
                           trace = TRUE)
#I just like to know how long this takes
end_time<-Sys.time()
end_time - start_time

#Inspect results
summary(multisession_fit_all)


```






Perhaps lets mess around with simulating multisession data?
```{r}
#What about multiple sessions
trap<-make.grid(detector = c("multi", rep ("proximity", 4)))
markocc(trap)<-c(1,0,0,0,0)


simMRCH_multi<-sim.capthist(trap, popn = list(D = 20, buffer = 100,
    Ndist = "poisson"), detectfn = 0, detectpar = list(),
    noccasions = 5, nsessions = 2, renumber = FALSE)

summary(simMRCH_multi, terse = TRUE)

sim.fit.multi<-secr.fit(simMRCH_multi, buffer = 500,trace = TRUE)
summary(sim.fit.multi)
predict(sim.fit.multi)


```














OLD
##############################



Multisession model
```{r}

#Create a vector of filenames for the traphistory
camera_names = c(file.path(input_directory, "deer_cameras_2018.txt"),
                 file.path(input_directory, "deer_cameras_2020.txt"),
                 file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_multisession.txt"), #This is all of the captures
              
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
#Summary by session
summary(deer_multisession)
#Summarised summary
summary(deer_multisession, terse = TRUE)

#Creates a series of plots, one for each session
plot(deer_multisession)

#Lets fit a preliminary model
multisession_fit<-secr.fit(deer_multisession, 
                           model = D~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 3),
                           buffer = 1000,trace = TRUE)
multisession_fit
coef(multisession_fit)


#Lets try something more complicated?!
library(MASS)
start_time<-Sys.time()
multisession_fit2<-secr.fit(deer_multisession, 
                           model = D~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 3,
                                          contrasts = list(session = MASS::contr.sdif)),
                           buffer = 1000,
                           trace = TRUE)
end_time<-Sys.time()
print(end_time - start_time)
coef(multisession_fit2)


#Ok so both of those above are giving me "zero" estimates for densities...lets try reducing the dataframe to two sessions and see what happens
#Just reduce everything down to 2021 and 2022
#Create a vector of filenames for the traphistory
camera_names = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_reduced.txt"), #This is all of the captures
              
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession, terse = TRUE)

multisession_fit<-secr.fit(deer_multisession, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2),
                           buffer = 300,
                           trace = TRUE)

multisession_fit<-secr.fit(deer_multisession, 
                          # model = D~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2),
                           buffer = 1000,
                           trace = TRUE)

multisession_fit2<-secr.fit(deer_multisession, 
                           model = D~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2,
                                          contrasts = list(session = MASS::contr.sdif)),
                           buffer = 1000,
                           trace = TRUE)

summary(multisession_fit2)

```




Going to try simulating stuff here to see whats happening
```{r}

grid<-make.grid(detector = c("multi", rep ("proximity", 4)))
markocc(grid)<-c(1,0,0,0,0)
g0mr<-c(0.3, 0.1, 0.1, 0.1, 0.1)
simMRCH<-sim.resight(grid, detectpar = list(g0 = g0mr, sigma = 25),
                     popn = list(D = 30), pID = 0.7)

summary(simMRCH)

sim.fit<-secr.fit(simMRCH, buffer = 500,trace = TRUE)
summary(sim.fit)


#Lets try to create a traps object
trap<-make.grid(detector = c("multi", rep ("proximity", 4)))
markocc(trap)<-c(1,0,0,0,0)


simMRCH<-sim.capthist(trap, popn = list(D = 5, buffer = 100,
    Ndist = "poisson"), detectfn = 0, detectpar = list(),
    noccasions = 5, nsessions = 1, binomN = NULL, exactN = NULL,
    p.available = 1, renumber = TRUE, seed = NULL,
    maxperpoly = 100, chulltol = 0.001, userdist = NULL, 
    savepopn = FALSE)

summary(simMRCH)

sim.fit<-secr.fit(simMRCH, buffer = 500,trace = TRUE)
summary(sim.fit)


#What about multiple sessions
trap<-make.grid(detector = c("multi", rep ("proximity", 4)))
markocc(trap)<-c(1,0,0,0,0)


simMRCH_multi<-sim.capthist(trap, popn = list(D = 5, buffer = 100,
    Ndist = "poisson"), detectfn = 0, detectpar = list(),
    noccasions = 5, nsessions = 2, binomN = NULL, exactN = NULL,
    p.available = 1, 
    renumber = FALSE, #Change this for multisessions
    seed = NULL,
    maxperpoly = 100, chulltol = 0.001, userdist = NULL, 
    savepopn = FALSE)

summary(simMRCH_multi)

sim.fit.multi<-secr.fit(simMRCH_multi, buffer = 500,trace = TRUE)
summary(sim.fit.multi)

#Ok now try to get two different density estimates, per session
sim.fit.multi<-secr.fit(simMRCH_multi, buffer = 500,trace = TRUE,
                        model = D~session)
summary(sim.fit.multi)

```

So maybe my multi session model is working? I just need to play with different starting values and let it run out?
```{r}
camera_names = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_multisession.txt"), #This is all of the captures
              
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession, terse = TRUE)

start_time<-Sys.time()
multisession_fit<-secr.fit(deer_multisession, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2),
                           buffer = 300,
                           trace = TRUE)
end_time<-Sys.time()
end_time - start_time

summary(multisession_fit)
predict(multisession_fit)


#Now lets try adding a list of unmarked sightings?

unmarked_obs = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

```

Lets try another simulation and see if I can recover those values
```{r}

grid<-make.grid(detector = c("multi", rep("proximity", 4)))
markocc(grid)<-c(1,0,0,0,0) #Marking occassions
g0mr<-c(0.3, 0.1, 0.1, 0.1, 0.1) #different detection rates on marking and sighting

simMRCH<-sim.resight(grid, detectpar=list(g0 = g0mr, sigma= 25),
                     popn = list(D = 30), pID = 0.7, seed = 42069)

sim.test<-secr.fit(simMRCH)
summary(sim.test)


#An example with sighting only data
grid<-make.grid(detector = "proximity")
markocc(grid)<-c(0,0,0,0,0)
plot(grid)

sim.test2<-sim.resight(grid, detectpar= list(g0 = 0.3, sigma = 25), unsighted = TRUE, #you can specify here that there are all zero capture histories
            popn = list(D = 30), pID = 1, seed = 123)
Tm(sim.test2)<-NULL
summary(sim.test2)

fit0<-secr.fit(sim.test2, fixed = list(pID = 1.0), 
         details = list(knownmarks = TRUE),
         trace = TRUE)

summary(fit0)


#Lets modify this to a multi session?
grid<-make.grid(detector = "count")
markocc(grid)<-c(0,0,0,0,0)
plot(grid)

sim.test3<-sim.capthist(grid, detectpar= list(g0 = 0.3, sigma = 25),
                       #nsessions = 1,
                      # unsighted = TRUE, #you can specify here that there are all zero capture histories
            popn = list(D = 30), seed = 123)
Tm(sim.test3)<-NULL
summary(sim.test3)

sim.test3<-sim.capthist(grid, popn = list(D = 30, buffer = 100, seed = 123),
             noccasions = 5,
             nsessions = 1)


fit0<-secr.fit(sim.test3, buffer = 100,trace = TRUE)

summary(fit0)

#Lets try to get estimates better
grid<-make.grid(detector = "count")
markocc(grid)<-c(rep(0, 10))
plot(grid)

sim.test4<-sim.capthist(grid, detectpar= list(g0 = 0.3, sigma = 25),
            popn = list(D = 30), 
            noccasions = 10,
            seed = 123)
summary(sim.test4)
fit0<-secr.fit(sim.test4, buffer = 500, 
               trace = TRUE)
fit4 <- secr.fit(sim.test4, buffer = 500, trace = TRUE, #this lines repeats the model specification from fit1
                 start = fit0, details = list(nsim = 10000)) #this line gives the previous model fit1 as a starting point, calls nsim simulations to estimated overdispersion (c-hat)


fit4<-secr.fit(sim.test4, buffer = 500, start = fit0,
        # details = list(nsim = 1000),
         trace = TRUE)
summary(fit4)
summary(fit0)


#Lets modify this to a multi session?#######
grid<-make.grid(detector = "count")
markocc(grid)<-c(1,0,0,0,0)
plot(grid)

sim.test3<-sim.capthist(grid, detectpar= list(g0 = 0.3, sigma = 25),
                       nsessions = 2, renumber = FALSE,
                      # unsighted = TRUE, #you can specify here that there are all zero capture histories
            popn = list(D = 30), seed = 123)
Tm(sim.test3)<-NULL
summary(sim.test3)



fit0<-secr.fit(sim.test3, 
               model = D~session,
               buffer = 500,trace = TRUE)

summary(fit0)

```


```{r}
#Trying it now with just sighting data
grid<-make.grid(detector = "count")
markocc(grid)<-c(0,0,0,0,0)

sim.test4<-sim.capthist(grid, detectpar= list(g0 = 0.3, sigma = 25),
                       nsessions = 2, renumber = FALSE,
            popn = list(D = 30), seed = 123)

summary(sim.test4)

fit1<-secr.fit(sim.test4, 
               model = D~session,
               buffer = 500,trace = TRUE)

summary(fit1)

browseURL("https://www.youtube.com/watch?v=JkrOQasshlY&list=RDJkrOQasshlY&start_radio=1")

############################################################
#Simulating data with a single marking occassion
grid<-make.grid(detector = "count")
markocc(grid)<-c(1,0,0,0,0)
#
sim.test3<-sim.capthist(grid, detectpar= list(g0 = 0.3, sigma = 25),
                       nsessions = 2, renumber = FALSE,
            popn = list(D = 30), seed = 123)


fit0<-secr.fit(sim.test3, 
               model = D~session,
               buffer = 500,trace = TRUE)

summary(fit0)

browseURL("https://www.youtube.com/watch?v=JkrOQasshlY&list=RDJkrOQasshlY&start_radio=1")

```

How does sim.resight work?
```{r}
?sim.resight()

grid <- make.grid(detector = 'count')
markocc(grid) <- c(0, 0, 0, 0, 0)
sim.test5 <- sim.resight(grid, detectpar=list(g0 = 0.3, sigma = 25), 
                         unsighted = TRUE, popn = list(D = 30), 
                         pID = 1.0, seed = 123)

summary(sim.test5)

fit5<-secr.fit(sim.test5, 
               buffer = 500,trace = TRUE)

summary(fit5)

browseURL("https://www.youtube.com/watch?v=JkrOQasshlY&list=RDJkrOQasshlY&start_radio=1")

```

Lets fit my multisession model and see whats going on
```{r}
#read in camera trap files - order matters here!
camera_names = c(file.path(input_directory, "deer_cameras_2021.txt"),
                 file.path(input_directory, "deer_cameras_2022.txt"))

#Prepare the capthist object
deer_multisession<-read.capthist(captfile = file.path(input_directory, "deer_reduced.txt"), 
              trapfile = camera_names,
              detector = "count",
              markocc = rep(0,30),
              skip = 1,
              verify = TRUE)
summary(deer_multisession, terse = TRUE)

start_time<-Sys.time()
#Fit the model
multisession_fit<-secr.fit(deer_multisession, 
                           model = D ~ session, #lowercase "session" to fit a value for each level of session
                           details = list(autoini = 2), #Was having getting started unless I specify
                           buffer = 500,
                           trace = TRUE)
#I just like to know how long this takes
end_time<-Sys.time()
end_time - start_time

#Inspect results
summary(multisession_fit)


```