The purpose of this script is to play around with secr models using known information on who is and isn't available in the populaiton

Setting up packages and directories
```{r}
#Just incase I wanted to know the run time of the whole thing!
start_time<-Sys.time()


#Clear everything out, its probably messy from the data prep file
rm(list = ls())

#Packages we will need
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(ggplot2)      #data visualization
#library(ggpubr)       #Arranging figures
library(stringr)      #reading information on strings
library(secr)         #the actual model engine
library(ggtext)       #Superscript on ggplot axis title
library(sf)           #Spatial data

#Set the working directories to read in the files we need
input_directory<-reproducible::checkPath(file.path(getwd(), "2. SECR Prep/outputs/multisession"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "3. SECR Models/multisession"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "3. SECR Models/multisession"), create = TRUE)
shapefiles_directory<-reproducible::checkPath(file.path(getwd(), "3. SECR Models/shapefiles"), create = TRUE)



```

2018 model specification. So I am going to do a couple things here
1. fit a regular 2018 model with just the baseline data.
2. fit a 2018 model with the additional information who is present but not sighted.
```{r}
#I will first have to read in some data I think and isolate for each session



deer_2018<-read.capthist(captfile = file.path(input_directory, "deer_2018.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2018.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2018, terse = TRUE)

#Lets look at the tracks
plot(deer_2018)

#Ok now try to add the unmarked sightings as you would normally? 
deer_2018<-addSightings(deer_2018, unmarked = file.path(input_directory, "unmarked_2018.txt"), skip = 1)

#I am going to use the original 2018 mask with a sufficiently large buffer, as that should encapsulate all the traps in remaining years despite small adjustments to trap locations. 
oak_bay_clipped_2018<-st_read(file.path(shapefiles_directory, "cameras_2018_clipped.shp"))
#Use that shape file to create a habitat mask for the initial model
mask_2018<-make.mask(traps(deer_2018), 
                type = "trapbuffer", 
                buffer = 1000, 
                poly = oak_bay_clipped_2018) #UPDATE THIS

#Fit the model - using autoini = 5 for starting values
fit_2018<-secr.fit(deer_2018, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2018,
                           details = list(autoini = 1), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2018)

#Save and extract results
mean_2018_unknown<-summary(fit_2018)[["predicted"]][["estimate"]][[1]]*100
upper_2018_unknown<-summary(fit_2018)[["predicted"]][["ucl"]][[1]]*100
lower_2018_unknown<-summary(fit_2018)[["predicted"]][["lcl"]][[1]]*100

##########################################################################
#Ok cool, now we try it with the known marks?
deer_2018_known<-read.capthist(captfile = file.path(input_directory, "deer_2018_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2018.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2018_known, terse = TRUE)

#Lets look at the tracks
plot(deer_2018_known)

#Ok now try to add the unmarked sightings as you would normally? 
deer_2018_known<-addSightings(deer_2018_known, unmarked = file.path(input_directory, "unmarked_2018.txt"), skip = 1)

#I am going to use the original 2018 mask with a sufficiently large buffer, as that should encapsulate all the traps in remaining years despite small adjustments to trap locations. 
oak_bay_clipped_2018<-st_read(file.path(shapefiles_directory, "cameras_2018_clipped.shp"))
#Use that shape file to create a habitat mask for the initial model
mask_2018<-make.mask(traps(deer_2018_known), 
                type = "trapbuffer", 
                buffer = 1000, 
                poly = oak_bay_clipped_2018) #UPDATE THIS

#Fit the model - using autoini = 5 for starting values
#fit_2018_known<-secr.fit(deer_2018_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2018,
                           details = list(autoini = 1), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
#summary(fit_2018_known)

#mean_2018_known<-summary(fit_2018_known)[["predicted"]][["estimate"]][[1]]*100
#upper_2018_known<-summary(fit_2018_known)[["predicted"]][["ucl"]][[1]]*100
#lower_2018_known<-summary(fit_2018_known)[["predicted"]][["lcl"]][[1]]*100


######################################
#Ok now the same thing but with the details argument...
#Ok cool, now we try it with the known marks?


#Fit the model - using autoini = 5 for starting values
fit_2018_known<-secr.fit(deer_2018_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2018,
                           details = list(autoini = 1,#Was having getting started unless I specify
                                          known = TRUE), #Specifying that the known marks are known
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

mean_2018_known<-summary(fit_2018_known)[["predicted"]][["estimate"]][[1]]*100
upper_2018_known<-summary(fit_2018_known)[["predicted"]][["ucl"]][[1]]*100
lower_2018_known<-summary(fit_2018_known)[["predicted"]][["lcl"]][[1]]*100

#CONCLUSION - the details(known = true) thing is default and important. 

#Ok lets carry on with the regularly scheduled programming. We need density estimates seperately for each year.



```

2019 model unknown and knonw
```{r}
deer_2019<-read.capthist(captfile = file.path(input_directory, "deer_2019.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2019.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2019, terse = TRUE)

#Lets look at the tracks
plot(deer_2019)

#Ok now try to add the unmarked sightings as you would normally? 
deer_2019<-addSightings(deer_2019, unmarked = file.path(input_directory, "unmarked_2019.txt"), skip = 1)

#I am going to use the original 2018 mask with a sufficiently large buffer, as that should encapsulate all the traps in remaining years despite small adjustments to trap locations. 
oak_bay_clipped_2018<-st_read(file.path(shapefiles_directory, "cameras_2018_clipped.shp"))
#Use that shape file to create a habitat mask for the initial model
mask_2019<-make.mask(traps(deer_2019), 
                type = "trapbuffer", 
                buffer = 1000, 
                poly = oak_bay_clipped_2018) #UPDATE THIS

#Fit the model - using autoini = 5 for starting values
fit_2019<-secr.fit(deer_2019, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2019)

#Save and extract results
mean_2019_unknown<-summary(fit_2019)[["predicted"]][["estimate"]][[1]]*100
upper_2019_unknown<-summary(fit_2019)[["predicted"]][["ucl"]][[1]]*100
lower_2019_unknown<-summary(fit_2019)[["predicted"]][["lcl"]][[1]]*100

#Also just checking something here, do we need to specify marks known or unknown?
#Unknown is default!!!!
# fit_2019_unknown<-secr.fit(deer_2019, 
#                           # model = D ~ session, #lowercase "session" to fit a value for each level of session
#                            mask = mask_2019,
#                            details = list(autoini = 1,
#                                           known = FALSE), #Was having getting started unless I specify
#                            buffer = 1000,
#                            trace = TRUE,
#                            detectfn = "HN", #Half normal detection function
#                             binomN = 0)  #counts modelled as poisson )
# 
# mean_2019_unknown<-summary(fit_2019_unknown)[["predicted"]][["estimate"]][[1]]*100
# upper_2019_unknown<-summary(fit_2019_unknown)[["predicted"]][["ucl"]][[1]]*100
# lower_2019_unknown<-summary(fit_2019_unknown)[["predicted"]][["lcl"]][[1]]*100


#Ok now lets fit with the known
deer_2019_known<-read.capthist(captfile = file.path(input_directory, "deer_2019_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2019.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2019_known, terse = TRUE)


#Ok now try to add the unmarked sightings as you would normally? 
deer_2019_known<-addSightings(deer_2019_known, unmarked = file.path(input_directory, "unmarked_2019.txt"), skip = 1)

#I am going to use the original 2018 mask with a sufficiently large buffer, as that should encapsulate all the traps in remaining years despite small adjustments to trap locations. 
oak_bay_clipped_2018<-st_read(file.path(shapefiles_directory, "cameras_2018_clipped.shp"))
#Use that shape file to create a habitat mask for the initial model
mask_2019<-make.mask(traps(deer_2019_known), 
                type = "trapbuffer", 
                buffer = 1000, 
                poly = oak_bay_clipped_2018) #UPDATE THIS

#Fit the model - using autoini = 5 for starting values
fit_2019_known<-secr.fit(deer_2019_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = TRUE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2019_known)

#Save and extract results
mean_2019_known<-summary(fit_2019_known)[["predicted"]][["estimate"]][[1]]*100
upper_2019_known<-summary(fit_2019_known)[["predicted"]][["ucl"]][[1]]*100
lower_2019_known<-summary(fit_2019_known)[["predicted"]][["lcl"]][[1]]*100


```


2020 model specifications
```{r}
deer_2020<-read.capthist(captfile = file.path(input_directory, "deer_2020.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2020.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2020, terse = TRUE)
deer_2020<-addSightings(deer_2020, unmarked = file.path(input_directory, "unmarked_2020.txt"), skip = 1)
#Fit the model - using autoini = 5 for starting values
fit_2020_unknown<-secr.fit(deer_2020, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 2,
                                          known = FALSE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Save and extract results
mean_2020_unknown<-summary(fit_2020_unknown)[["predicted"]][["estimate"]][[1]]*100
upper_2020_unknown<-summary(fit_2020_unknown)[["predicted"]][["ucl"]][[1]]*100
lower_2020_unknown<-summary(fit_2020_unknown)[["predicted"]][["lcl"]][[1]]*100
#########################################################################
deer_2020_known<-read.capthist(captfile = file.path(input_directory, "deer_2020_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2020.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2020_known, terse = TRUE)
deer_2020_known<-addSightings(deer_2020_known, unmarked = file.path(input_directory, "unmarked_2020.txt"), skip = 1)
fit_2020_known<-secr.fit(deer_2020_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(#autoini = 1,
                                          known = TRUE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )



#Inspect those initial results
summary(fit_2020_known)

#Save and extract results
mean_2020_known<-summary(fit_2020_known)[["predicted"]][["estimate"]][[1]]*100
upper_2020_known<-summary(fit_2020_known)[["predicted"]][["ucl"]][[1]]*100
lower_2020_known<-summary(fit_2020_known)[["predicted"]][["lcl"]][[1]]*100

```


2021 model specification
```{r}
deer_2021<-read.capthist(captfile = file.path(input_directory, "deer_2021.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2021.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2021, terse = TRUE)
deer_2021<-addSightings(deer_2021, unmarked = file.path(input_directory, "unmarked_2021.txt"), skip = 1)
#Fit the model - using autoini = 5 for starting values
fit_2021_unknown<-secr.fit(deer_2021, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = FALSE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Save and extract results
mean_2021_unknown<-summary(fit_2021_unknown)[["predicted"]][["estimate"]][[1]]*100
upper_2021_unknown<-summary(fit_2021_unknown)[["predicted"]][["ucl"]][[1]]*100
lower_2021_unknown<-summary(fit_2021_unknown)[["predicted"]][["lcl"]][[1]]*100
#########################################################################
deer_2021_known<-read.capthist(captfile = file.path(input_directory, "deer_2021_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2021.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2021_known, terse = TRUE)
deer_2021_known<-addSightings(deer_2021_known, unmarked = file.path(input_directory, "unmarked_2021.txt"), skip = 1)
fit_2021_known<-secr.fit(deer_2021_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = TRUE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2021_known)

#Save and extract results
mean_2021_known<-summary(fit_2021_known)[["predicted"]][["estimate"]][[1]]*100
upper_2021_known<-summary(fit_2021_known)[["predicted"]][["ucl"]][[1]]*100
lower_2021_known<-summary(fit_2021_known)[["predicted"]][["lcl"]][[1]]*100

```


2022 model specification
```{r}
deer_2022<-read.capthist(captfile = file.path(input_directory, "deer_2022.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2022.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2022, terse = TRUE)
deer_2022<-addSightings(deer_2022, unmarked = file.path(input_directory, "unmarked_2022.txt"), skip = 1)
#Fit the model - using autoini = 5 for starting values
fit_2022_unknown<-secr.fit(deer_2022, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = FALSE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Save and extract results
mean_2022_unknown<-summary(fit_2022_unknown)[["predicted"]][["estimate"]][[1]]*100
upper_2022_unknown<-summary(fit_2022_unknown)[["predicted"]][["ucl"]][[1]]*100
lower_2022_unknown<-summary(fit_2022_unknown)[["predicted"]][["lcl"]][[1]]*100
#########################################################################
deer_2022_known<-read.capthist(captfile = file.path(input_directory, "deer_2022_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2022.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2022_known, terse = TRUE)
deer_2022_known<-addSightings(deer_2022_known, unmarked = file.path(input_directory, "unmarked_2022.txt"), skip = 1)
fit_2022_known<-secr.fit(deer_2022_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = TRUE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2022_known)

#Save and extract results
mean_2022_known<-summary(fit_2022_known)[["predicted"]][["estimate"]][[1]]*100
upper_2022_known<-summary(fit_2022_known)[["predicted"]][["ucl"]][[1]]*100
lower_2022_known<-summary(fit_2022_known)[["predicted"]][["lcl"]][[1]]*100

```


2023 model specification
```{r}
deer_2023<-read.capthist(captfile = file.path(input_directory, "deer_2023.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2023.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2023, terse = TRUE)
deer_2023<-addSightings(deer_2023, unmarked = file.path(input_directory, "unmarked_2023.txt"), skip = 1)
#Fit the model - using autoini = 5 for starting values
fit_2023_unknown<-secr.fit(deer_2023, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = FALSE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Save and extract results
mean_2023_unknown<-summary(fit_2023_unknown)[["predicted"]][["estimate"]][[1]]*100
upper_2023_unknown<-summary(fit_2023_unknown)[["predicted"]][["ucl"]][[1]]*100
lower_2023_unknown<-summary(fit_2023_unknown)[["predicted"]][["lcl"]][[1]]*100
#########################################################################
deer_2023_known<-read.capthist(captfile = file.path(input_directory, "deer_2023_known.txt"), 
              trapfile = file.path(input_directory, "deer_cameras_2023.txt"), #A list of files of camera locations for each year
              detector = "count",
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1,
              verify = TRUE)
summary(deer_2023_known, terse = TRUE)
deer_2023_known<-addSightings(deer_2023_known, unmarked = file.path(input_directory, "unmarked_2023.txt"), skip = 1)
fit_2023_known<-secr.fit(deer_2023_known, 
                          # model = D ~ session, #lowercase "session" to fit a value for each level of session
                           mask = mask_2019,
                           details = list(autoini = 1,
                                          known = TRUE), #Was having getting started unless I specify
                           buffer = 1000,
                           trace = TRUE,
                           detectfn = "HN", #Half normal detection function
                            binomN = 0)  #counts modelled as poisson )

#Inspect those initial results
summary(fit_2023_known)

#Save and extract results
mean_2023_known<-summary(fit_2023_known)[["predicted"]][["estimate"]][[1]]*100
upper_2023_known<-summary(fit_2023_known)[["predicted"]][["ucl"]][[1]]*100
lower_2023_known<-summary(fit_2023_known)[["predicted"]][["lcl"]][[1]]*100

```

And now a chunk to plot all those results!
```{r}
#Wrap that into a dataframe and plot
density_estimates<-data.frame(
    year = c(2018, 2019, 2020, 2021, 2022, 2023),
    mean_density = c(mean_2018, mean_2019, mean_2020, mean_2021, mean_2022, mean_2023),
    upper_cl = c(upper_2018, upper_2019, upper_2020, upper_2021, upper_2022, upper_2023),
    lower_cl = c(lower_2018, lower_2019, lower_2020, lower_2021, lower_2022, lower_2023)
)

density_estimates<-rbind(data.frame(year = c(2018, 2019, 2020, 2021, 2022, 2023),
                              type = c("unknown","unknown","unknown","unknown","unknown","unknown"),
                              mean_density = c(mean_2018_unknown,
                                               mean_2019_unknown,
                                               mean_2020_unknown,
                                               mean_2021_unknown,
                                               mean_2022_unknown,
                                               mean_2023_unknown),
                              upper_ci = c(upper_2018_unknown,
                                           upper_2019_unknown,
                                           upper_2020_unknown,
                                           upper_2021_unknown,
                                           upper_2022_unknown,
                                           upper_2023_unknown),
                              lower_ci = c(lower_2018_unknown,
                                           lower_2019_unknown,
                                           lower_2020_unknown,
                                           lower_2021_unknown,
                                           lower_2022_unknown,
                                           lower_2023_unknown)),
data.frame(year = c(2018, 2019, # 2020, 
                    2021, 2022, 2023),
                              type = c("known","known","known","known","known"),
                              mean_density = c(mean_2018_known,
                                               mean_2019_known,
                                               #mean_2020_known,
                                               mean_2021_known,
                                               mean_2022_known,
                                               mean_2023_known),
                              upper_ci = c(upper_2018_known,
                                           upper_2019_known,
                                          # upper_2020_known,
                                           upper_2021_known,
                                           upper_2022_known,
                                           upper_2023_known),
                              lower_ci = c(lower_2018_known,
                                           lower_2019_known,
                                          # lower_2020_known,
                                           lower_2021_known,
                                           lower_2022_known,
                                           lower_2023_known)))

ggplot(density_estimates, aes(x = as.factor(year), y = mean_density, group = type)) +
  geom_point(aes(colour = type), position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci, colour = type), 
                width = 0.2, position = position_dodge(width = 0.5)) +
  ylab("Mean Density ± 95% CI") +
  xlab("Year") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 12))
                              


```