The purpose of this script is to analyze detection data from an agent-based model of deer movement in Oak Bay, British Columbia. This script relies on data from urban_deer_movement_simulation.nlogo, located in the sibling folder 5.1 Netlogo Model. This script produces a series of single session mark resight models in a similar framework to the main paper's analysis. The main outputs of this script are density estimates for each simulated scenario, and a figure comparing density estimates using male & female data vs just female deer data.

Andrew Barnas
andrewbarnas@uvic.ca
April 9th 2026


Load packages and data
```{r}

rm(list=ls())
#Being a bit lazy here, setting working directory to file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

#Packages needed
library(dplyr)        #data wrangling
library(tidyr)        #more data wrangling
library(purrr)        #feline friendly loops
library(stringr)      #dealing with strings
library(readr)
library(secr)         #density estimates
library(reproducible) #reproducible workflows
library(ggplot2)      #data visualization

#Setup working directories
netlogo_directory<-reproducible::checkPath(file.path(getwd(), "5.1 Netlogo Model/outputs"), create = TRUE)
input_directory<-reproducible::checkPath(file.path(getwd(), "inputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "figures"), create = TRUE)
```


Ok all I want to do with this block is read in the camera operability files and the deteciton files into lists
```{r}
list.files(netlogo_directory)

# 1. Identify the files
camop_paths <- list.files(netlogo_directory, 
                          pattern = "camera_operability_.*\\.csv$", 
                          full.names = TRUE)

#Read them into a giant dataframe
camop_df<-read_csv(camop_paths)

##################################################
#Now just need to calculate usage, this can be done safely as it is done by row
# --- 1. Setup Constants ---
patch_size_m  <- 11.18
utm_x_origin  <- 476000
utm_y_origin  <- 5362000

camop_df<-camop_df%>%
  mutate(
      trap.id = camera_id,
      # Convert patch coords -> meters -> UTM
      x = round(utm_x_origin + (xcor * patch_size_m), 0),
      y = round(utm_y_origin + (ycor * patch_size_m), 0),
      # Calculate usage string
      n_days = ticks / (60 * 24),
      usage = str_dup("1", n_days)
    ) %>%
    select(run_num, number_deer, trap.id, x, y, usage) %>%
    arrange(run_num, number_deer, trap.id)

#Nice, and for subsequent moves we can just pull out the run number of interest!
camop_df%>%
  filter(run_num == 3)


```

Ok now the deteciton files. Same thing these need to be read into a giant dataframe that we can subsequently pull from
```{r}
# 1. Identify the files
deer_paths <- list.files(netlogo_directory, 
                          pattern = "deer_detections_.*\\.csv$", 
                          full.names = TRUE)

#Read them into a giant dataframe
deer_df<-read_csv(deer_paths)


#One quick rename and re-arrange for aesthetics
deer_df<-deer_df%>%
  rename(number_deer = `number-deer`)%>%
  relocate(run_num, number_deer)%>%
  arrange(run_num, number_deer)

```


Ok so I have a bunch of files of deer detections, and as of right now they are all of "known" IDs. But we need to make these into subsets of marked and unmarked individuals to replicate reality. But, I want to make sure that the deer I randomly assign to be "known" within each simulation is consistent right, to make sure we can replicate results.
```{r}

#So what I am going to do here is go dataframe by dataframe, and pull out the same random sample of individuals to be marked, label them as marked, and add those dataframes to a growing list that we can work from!

#Start with an empty list
dataframes<-list()

for(i in unique(deer_df$run_num)){
  
  #Pull out the dataframe
  temp<-deer_df%>%filter(run_num == i)
  
  #Number to sample
  number_to_sample<-as.numeric(unique(temp$number_deer))/4
  
  #Figure out who the marked individuals should be
  marked_female_ids<-temp%>%
    filter(sex == "female")%>%
    distinct(deer_id)%>%
    pull(deer_id)%>%
    sample(number_to_sample)
  
  #And then assign status based on whether they occur in the marked dataframe
  temp<-temp%>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))
  
  #And then add the dataframe to the growing list
  dataframes <- append(dataframes, list(temp))
  
}


#Ok, now we have a way to pull out data for specific run numbers from our list of dataframes, and the camop dataframe!
#For example pulling out runnumber 2

# Find the index of the list element with run_num == X
index <- which(sapply(dataframes, function(df) unique(df$run_num) == 2))

# Extract the corresponding dataframe
deer <- dataframes[[index]]
camop<-camop_df%>%
  filter(run_num == 2)

```




##############################################################################################
Everything past here I am going to deal with seperately our of laziness/ease. For each of the scenarios I need to 1) identify indepenent events for marked and unmarked individuals, 2) format those files for secr, 3) write those files to text

#################################################################################################
######### POPULATION SIZE: 400 (NO REDUCTION)###################################################
#################################################################################################
```{r}


#Ok so we have two things here, a list of deer detection dataframes and a giant dataframe of camop stuff
dataframes
camop_df

#In the following loop we are going to work through them by each unique run number by subsetting out the data of interest
#And carrying on like we did previously


#A list to store the unmarked matrix counts
unmarked_matrix_list<-list()
#A list to store counts of known individuals
known_individuals_list<-list()

for (j in unique(camop_df$run_num)) {
  
  #And here is the new procedure for pulling out dataframes of interest
  # Find the index of the list element with run_num == X
  index <- which(sapply(dataframes, function(df) unique(df$run_num) == j))
  df <- dataframes[[index]]
  
  camop<-camop_df%>%
  filter(run_num == j)
  
  
  
  
  
  df_unmarked<-df%>%
  filter(status == "unmarked")
 #And the marked dataframe, we will deal with this below
  df_marked<-df%>%
  filter(status == "marked" )
  
  #But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, run_num, number_deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)
 
  
  # Store the result in the i-th position of the list
  unmarked_matrix_list[[j]] <- unmarked_matrix
  
  
  ####################################################
  #Now the marked individuals
  
 df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)
 
 # Store the result in the i-th position of the list
  known_individuals_list[[j]] <- df_independent 
  
  
  #I've got a crazy idea here...since we know the population size and run number we are dealing with
  #Can we not just export these data on the fly?
  #To export
  df_independent
  unmarked_matrix
  camop%>%dplyr::select(-c(run_num, number_deer)) #modified
  
  #store the population size we will need it for naming
  num_deer<-as.numeric(unique(camop$number_deer))
  
  
  write.table(df_independent,
              file.path(input_directory, paste0("df_independent_", num_deer, "_run", j,  ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
  
  write.table(unmarked_matrix,
              file.path(input_directory, paste0("unmarked_matrix_", num_deer, "_run", j,  ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
   
  write.table(camop%>%dplyr::select(-c(run_num, number_deer)),
              file.path(input_directory, paste0("camop_", num_deer, "_run", j,  ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
  
}


#Extract unmarked counts
#unmarked_matrix_400<-unmarked_matrix_list[[1]]
#unmarked_matrix_360<-unmarked_matrix_list[[2]]
#unmarked_matrix_300<-unmarked_matrix_list[[3]]
#unmarked_matrix_200<-unmarked_matrix_list[[4]]

#extract known counts
#df_independent_400<-known_individuals_list[[1]]
#df_independent_360<-known_individuals_list[[2]]
#df_independent_300<-known_individuals_list[[3]]
#df_independent_200<-known_individuals_list[[4]]



```



Now just females!!
#################################################################################################
######### FEMALES ONLY!!!########################################################################
#################################################################################################
```{r}

#Same thing as above still using the same dataframes and lists



#A list to store the unmarked matrix counts
unmarked_matrix_females_list<-list()
#A list to store counts of known individuals
known_individuals_females_list<-list()

for (j in unique(camop_df$run_num)) {
  
  #And here is the new procedure for pulling out dataframes of interest
  # Find the index of the list element with run_num == X
  index <- which(sapply(dataframes, function(df) unique(df$run_num) == j))
  df <- dataframes[[index]]
  
  camop<-camop_df%>%
  filter(run_num == j)
  
  
  #KEY DIFFERENCE HERE! ONLY RETAINING FEMALE DATA
  df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
 #And the marked dataframe, we will deal with this below
  df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female")
  
  #But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, run_num, number_deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)
 
  
  # Store the result in the i-th position of the list
  unmarked_matrix_females_list[[j]] <- unmarked_matrix
  
  
  ####################################################
  #Now the marked individuals
  
 df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)
 
 # Store the result in the i-th position of the list
  known_individuals_females_list[[j]] <- df_independent 
  
  #Replicating my crazy idea of outputing on the go
   #store the population size we will need it for naming
  num_deer<-as.numeric(unique(camop$number_deer))
  
  
  write.table(df_independent,
              file.path(input_directory, paste0("df_independent_", num_deer, "_run", j, "_females",  ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
  
  write.table(unmarked_matrix,
              file.path(input_directory, paste0("unmarked_matrix_", num_deer, "_run", j, "_females",  ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
   
  write.table(camop%>%dplyr::select(-c(run_num, number_deer)),
              file.path(input_directory, paste0("camop_", num_deer, "_run", j, "_females", ".txt")), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
  
  
  
  
}


```


Ok we are almost there. Now I have a directory filed with files and I need to loop over the right ones



And now we can fit models for all
```{r}

list.files(input_directory)

# Setup your variables
pop_sizes <- c("400", "360", "300", "200")
groups <- c("all", "females")
runs<-paste0("run", unique(camop_df$run_num))


# Outer loop for population size
for (size in pop_sizes) {
  
  # Inner loop for group (All vs Females)
  for (group in groups) {
    
    #And then finally the run number of interest
    for (run in runs){
    
    # Construct file names dynamically
    # Note: Using the naming logic from your list.files()
    suffix <- ifelse(group == "females", "_females", "")
    
    capt_file <- file.path(input_directory, paste0("df_independent_", size, "_", run, suffix, ".txt"))
    trap_file <- file.path(input_directory, paste0("camop_", size, "_", run, suffix, ".txt"))
    unmarked_file <- file.path(input_directory,paste0("unmarked_matrix_", size, "_", run, suffix, ".txt"))
    
    
    # Optional: skip if files don't exist (VERY useful here)
      if (!all(file.exists(c(capt_file, trap_file, unmarked_file)))) {
        message("Skipping missing combo: ", size, " ", group, " ", run)
        next
      }
    
    # 1. Read capthist
    deer_capthist <- read.capthist(
      captfile = capt_file, 
      trapfile = trap_file,
      detector = "count",
      markocc = rep(0, 30),
      skip = 1,
      verify = TRUE
    )
    
    # 2. Add unmarked sightings
    deer_capthist <- addSightings(
      deer_capthist, 
      unmarked = unmarked_file, 
      skip = 1
    )
    
    # 3. Fit the model
    # I recommend adding a message so you know which model is running!
    message(paste("Fitting model for size:", size, "| group:", group, "|", run))
    
    fit_model <- secr.fit(
      deer_capthist, 
      buffer = 1000, 
      detectfn = "HN",
      binomN = 0,
      trace = FALSE # Set to FALSE in loops to keep console clean, or TRUE to monitor
    )
    
    # 4. Save the model object
    # We create a unique filename like "fit_deer_secr_400_females.RData"
    out_filename <- paste0("fit_deer_secr_", size, "_", group, "_", run, ".RData")
    save(fit_model, file = file.path(output_directory, out_filename))
    
    # Clean up memory after each fit
    rm(fit_model, deer_capthist)
    gc()
    
    }
  }
}




```



Bring the saved model objects back in, extract estimates and plot!
```{r}
pop_sizes <- c("300", "200")
groups <- c("all", "females")
runs<-paste0("run", unique(camop_df$run_num))

# Create an empty dataframe to store results
results <- data.frame()

for (size in pop_sizes) {
  for (group in groups) {
    for(run in runs){
    
    file_name <- paste0("fit_deer_secr_", size, "_", group, "_", run, ".RData")
    file_path <- file.path(output_directory, file_name)
    
    # Only proceed if the file actually exists
    if (file.exists(file_path)) {
      
      # Load the file. Remember, inside the loop it was saved as 'fit_model'
      load(file_path)
      
      # Extract predicted density (usually the first row of the 'predicted' table)
      stats <- summary(fit_model)$predicted
      
      # Create a temporary row
      temp_df <- data.frame(
        population_size = as.numeric(size),
        sex_group = group,
        run = run,
        density = stats["D", "estimate"] * 100,
        ucl = stats["D", "ucl"] * 100,
        lcl = stats["D", "lcl"] * 100
      )
      
      # Bind to the main results table
      results <- rbind(results, temp_df)
      
      # Clean up to keep the environment tidy
      rm(fit_model)
     }
   }
  }
}

#Clean up the dataframe a bit
results<-results %>%
  group_by(population_size, sex_group) %>%
  mutate(model_id = row_number()) %>%
  ungroup()

results$population_size<-factor(results$population_size,
                                   levels = c("300", "200"))


# 2. Plot with nested dodging
ggplot(results, aes(x = as.factor(population_size), 
                    y = density, 
                    color = sex_group,
                    # Putting sex_group last in interaction keeps colors together
                    group = interaction(model_id, sex_group))) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), 
                position = position_dodge(width = 0.3), 
                width = 0.2) +
  geom_point(position = position_dodge(width = 0.3), 
             size = 3) +
  labs(x = "Population Size", y = "Density Estimate", color = "Sex Group") +
  theme_classic()




#And plot
ggplot(results, aes(x = as.factor(population_size), 
                           y = density, 
                           color = sex_group, 
                           group = sex_group)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  geom_line(aes(group = sex_group), 
            position = position_dodge(width = 0.5), 
            linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), 
             size = 3) +
  # Add custom labels here:
  scale_x_discrete(labels = c("400" = "Original Population\n(N = 400)", 
                              "360" = "10% Reduction\n(N = 360)", 
                              "300" = "25% Reduction\n(N = 300)", 
                              "200" = "50% Reduction\n(N = 200)")) +
  labs(x = "Simulated Population Size", 
       y = "Estimated Density", 
       color = "Data Subset") +
  theme_minimal()

#And plot
ggplot(results, aes(x = as.factor(population_size), 
                           y = density, 
                           color = sex_group, 
                           group = sex_group)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  geom_line(aes(group = sex_group), 
            position = position_dodge(width = 0.5), 
            linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), 
             size = 3) +
  # Add custom labels here:
  scale_x_discrete(labels = c("400" = "Original Population\n(N = 400)", 
                              "360" = "10% Reduction\n(N = 360)", 
                              "300" = "25% Reduction\n(N = 300)", 
                              "200" = "50% Reduction\n(N = 200)")) +
  labs(x = "Simulated Population Size", 
       y = "Estimated Density", 
       color = "Data Subset") +
  theme_minimal()






```




