The purpose of this script is to analyze detection data from an agent-based model of deer movement in Oak Bay, British Columbia. This script relies on data from urban_deer_movement_simulation.nlogo, located in the sibling folder 5.1 Netlogo Model. This script produces a series of single session mark resight models in a similar framework to the main paper's analysis. The main outputs of this script are density estimates for each simulated scenario, and a figure comparing density estimates using male & female data vs just female deer data.

Andrew Barnas
andrewbarnas@uvic.ca
April 9th 2026


Load packages and data
```{r}

rm(list=ls())
#Being a bit lazy here, setting working directory to file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

#Packages needed
library(dplyr)        #data wrangling
library(tidyr)        #more data wrangling
library(stringr)      #dealing with strings
library(secr)         #density estimates
library(reproducible) #reproducible workflows
library(ggplot2)      #data visualization

#Setup working directories
input_directory<-reproducible::checkPath(file.path(getwd(), "5.1 Netlogo Model/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "5. Agent Based Simulation/5.2 Density Models/outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "5. Agent Based Simulation/5.2 Density Models/figures"), create = TRUE)


#Read in the datafiles we will need.
#Sepearate detection histories
df_400<-read.csv(file.path(input_directory, "deer_detections_400.csv"), header = TRUE)
df_360<-read.csv(file.path(input_directory, "deer_detections_360.csv"), header = TRUE)
df_300<-read.csv(file.path(input_directory, "deer_detections_300.csv"), header = TRUE)
df_200<-read.csv(file.path(input_directory, "deer_detections_200.csv"), header = TRUE)

#Camera operability
camop_400<-read.csv(file.path(input_directory, "camera_operability_400.csv"), header = TRUE)
camop_360<-read.csv(file.path(input_directory, "camera_operability_360.csv"), header = TRUE)
camop_300<-read.csv(file.path(input_directory, "camera_operability_300.csv"), header = TRUE)
camop_200<-read.csv(file.path(input_directory, "camera_operability_200.csv"), header = TRUE)


```

Ok so in the following sections I do a fair bit of copying and pasting, because 1) I am trying to meet a review deadline, 2) this is easier for now. If I had time I would fancify this, but do as I say not as I do.
But, there are some things I can accomplish once and use them throughout while keeping my environment clean
```{r}
#Patch size of netlogo simulations, used for space calculations
patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#Quickly create the camop_traps file for all scenarios (400, 360, 300, 200)
camop_400 <- camop_400 %>%
  mutate(
    trap.id = camera_id,
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    # Number of days
    n_days = ticks / (60 * 24),
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

camop_360 <- camop_360 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))


camop_300 <- camop_300 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))


camop_200 <- camop_200 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```

##############################################################################################
Everything past here I am going to deal with seperately our of laziness/ease. For each of the scenarios I need to 1) identify indepenent events for marked and unmarked individuals, 2) format those files for secr, 3) write those files to text

Ok lets pull out some proportion of individuals and make them marked vs unmarked. Need to introduce some error ya know
```{r}
#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(100)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```

Export those files to text (needed for secr)
```{r}
write.table(camop_traps, "camop.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent, "df_independent_400_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_400_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

########################################################################################
OKAY NOW THE SAME THING AS ABOVE BUT WE ARE ONLY RETAINING INFORMATION FOR FEMALE DEER THROUGHOUT
```{r}
#Read the original frame back in for safety
df<-read.csv("deer_detections_400.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(100)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female") #redundant but I like it

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```


Export those files to text (needed for secr)
```{r}

write.table(df_independent, "df_independent_400_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_400_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

Now lets fit models for both!
```{r}
#First the full model with all individuals

deer_secr_400_all<-read.capthist(captfile = file.path(getwd(), "df_independent_400_all.txt"), 
              trapfile = file.path(getwd(), "camop.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_400_all)
plot(deer_secr_400_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_400_all<-addSightings(deer_secr_400_all, unmarked = file.path(getwd(), "unmarked_matrix_400_all.txt"), skip = 1)

#Fit the model 
fit_deer_secr_400_all<-secr.fit(deer_secr_400_all, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_400_all)

save(fit_deer_secr_400_all, file = file.path(getwd(), "fit_deer_secr_400_all.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_400_all<-summary(fit_deer_secr_400_all)[["predicted"]][["estimate"]][[1]]*100
ucl_400_all<-summary(fit_deer_secr_400_all)[["predicted"]][["ucl"]][[1]]*100
lcl_400_all<-summary(fit_deer_secr_400_all)[["predicted"]][["lcl"]][[1]]*100
mean_400_all
ucl_400_all
lcl_400_all

###########################################################
#And now an estimate with just the females
deer_secr_400_females<-read.capthist(captfile = file.path(getwd(), "df_independent_400_females.txt"), 
              trapfile = file.path(getwd(), "camop.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_400_females)
plot(deer_secr_400_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_400_females<-addSightings(deer_secr_400_females, unmarked = file.path(getwd(), "unmarked_matrix_400_females.txt"), skip = 1)

#Fit the model 
fit_deer_secr_400_females<-secr.fit(deer_secr_400_females, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_400_females)

save(fit_deer_secr_400_females, file = file.path(getwd(), "fit_deer_secr_400_females.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_400_females<-summary(fit_deer_secr_400_females)[["predicted"]][["estimate"]][[1]]*100
ucl_400_females<-summary(fit_deer_secr_400_females)[["predicted"]][["ucl"]][[1]]*100
lcl_400_females<-summary(fit_deer_secr_400_females)[["predicted"]][["lcl"]][[1]]*100
mean_400_females
ucl_400_females
lcl_400_females


```


#######################################################################################################################
############################# REPEATING FOR LOWER POPULATION SIZE #####################################################
#######################################################################################################################

Need to re-do the camera operability stuff here because technically its a different run
```{r}
#read in the camop from the 200 run
camop<-read.csv(file.path(getwd(), "camera_operability_200.csv"), header = TRUE)

#This should be really easy since I am programming in perfect operability amongst all cameras. I have the model spit out a file of camera locations, and the number of ticks each camera was operating until

patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#And then transform the camop file to match the format we need:

camop_traps <- camop %>%
  mutate(
    trap.id = camera_id,
    
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    
    # Number of days
    n_days = ticks / (60 * 24),
    
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)

head(camop_traps)

camop_traps<-camop_traps%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```


Ok lets pull out some proportion of individuals and make them marked vs unmarked. Need to introduce some error ya know
```{r}
#read in the lower population size data
df<-read.csv("deer_detections_200.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
set.seed(420)
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(50)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```

Export those files to text (needed for secr)
```{r}
write.table(camop_traps, "camop_200.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent, "df_independent_200_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_200_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

########################################################################################
OKAY NOW THE SAME THING AS ABOVE BUT WE ARE ONLY RETAINING INFORMATION FOR FEMALE DEER THROUGHOUT
```{r}
#Read the original frame back in for safety
df<-read.csv("deer_detections_200.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
set.seed(420)
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(50)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female") #redundant but I like it

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```


Export those files to text (needed for secr)
```{r}
#Camop file is the same so no need to redo here

write.table(df_independent, "df_independent_200_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_200_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

Now lets fit models for both!
```{r}
#First the full model with all individuals

deer_secr_200_all<-read.capthist(captfile = file.path(getwd(), "df_independent_200_all.txt"), 
              trapfile = file.path(getwd(), "camop_200.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_200_all)
plot(deer_secr_200_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_200_all<-addSightings(deer_secr_200_all, unmarked = file.path(getwd(), "unmarked_matrix_200_all.txt"), skip = 1)

#Fit the model 
fit_deer_secr_200_all<-secr.fit(deer_secr_200_all, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_200_all)

save(fit_deer_secr_200_all, file = file.path(getwd(), "fit_deer_secr_200_all.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["estimate"]][[1]]*100
ucl_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["ucl"]][[1]]*100
lcl_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["lcl"]][[1]]*100
mean_200_all
ucl_200_all
lcl_200_all

###########################################################
#And now an estimate with just the females
deer_secr_200_females<-read.capthist(captfile = file.path(getwd(), "df_independent_200_females.txt"), 
              trapfile = file.path(getwd(), "camop_200.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_200_females)
plot(deer_secr_200_females, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_200_females<-addSightings(deer_secr_200_females, unmarked = file.path(getwd(), "unmarked_matrix_200_females.txt"), skip = 1)

#Fit the model 
fit_deer_secr_200_females<-secr.fit(deer_secr_200_females, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_200_females)

save(fit_deer_secr_200_females, file = file.path(getwd(), "fit_deer_secr_200_females.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["estimate"]][[1]]*100
ucl_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["ucl"]][[1]]*100
lcl_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["lcl"]][[1]]*100
mean_200_females
ucl_200_females
lcl_200_females


```


Lets wrap these into dataframes for plotting
```{r}

density_figure<-data.frame(mean_density = c(mean_400_all, mean_400_females, mean_200_all, mean_200_females),
           lc_density = c(lcl_400_all, lcl_400_females, lcl_200_all, lcl_200_females),
           uc_density = c(ucl_400_all, ucl_400_females, ucl_200_all, ucl_200_females),
           population = c("large", "large", "small", "small"),
           data = c("all", "only females", "all", "only females"))

ggplot(density_figure, aes(x = as.factor(population), y = mean_density))+
  geom_point()+
  geom_errorbar(aes(ymin = lc_density, ymax = uc_density))


ggplot(density_figure, aes(x = as.factor(population), y = mean_density, color = data)) +
  # Use position_dodge to separate "all" and "only females" horizontally
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  # The width inside position_dodge must match the geom_point to keep them aligned
  geom_errorbar(aes(ymin = lc_density, ymax = uc_density), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  labs(x = "Population Size", y = "Estimated Density", color = "Data Subset") +
  theme_minimal()

ggplot(density_figure, aes(x = as.factor(population), 
                           y = mean_density, 
                           color = data, 
                           group = data)) + # Grouping by 'data' connects the points
  geom_errorbar(aes(ymin = lc_density, ymax = uc_density), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  geom_line(aes(group = data), 
            position = position_dodge(width = 0.5), 
            linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), 
             size = 3) +
  labs(x = "Population Size", 
       y = "Estimated Density", 
       color = "Data Subset") +
  theme_minimal()

```

#################################################################################################
######### POPULATION SIZE: 360 (10% REDUCTION)###################################################
#################################################################################################

Need to re-do the camera operability stuff here because technically its a different run
```{r}
#read in the camop from the 200 run
camop<-read.csv(file.path(getwd(), "camera_operability_360.csv"), header = TRUE)

#This should be really easy since I am programming in perfect operability amongst all cameras. I have the model spit out a file of camera locations, and the number of ticks each camera was operating until

patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#And then transform the camop file to match the format we need:

camop_traps <- camop %>%
  mutate(
    trap.id = camera_id,
    
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    
    # Number of days
    n_days = ticks / (60 * 24),
    
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)

head(camop_traps)

camop_traps<-camop_traps%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```


Ok lets pull out some proportion of individuals and make them marked vs unmarked. Need to introduce some error ya know
```{r}
#read in the lower population size data
df<-read.csv("deer_detections_360.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(90)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```

Export those files to text (needed for secr)
```{r}
write.table(camop_traps, "camop_360.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent, "df_independent_360_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_360_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

########################################################################################
OKAY NOW THE SAME THING AS ABOVE BUT WE ARE ONLY RETAINING INFORMATION FOR FEMALE DEER THROUGHOUT
```{r}
#Read the original frame back in for safety
df<-read.csv("deer_detections_360.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(90)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female") #redundant but I like it

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```


Export those files to text (needed for secr)
```{r}
#Camop file is the same so no need to redo here

write.table(df_independent, "df_independent_360_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_360_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

Now lets fit models for both!
```{r}
#First the full model with all individuals

deer_secr_360_all<-read.capthist(captfile = file.path(getwd(), "df_independent_360_all.txt"), 
              trapfile = file.path(getwd(), "camop_360.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_360_all)
plot(deer_secr_360_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_360_all<-addSightings(deer_secr_360_all, unmarked = file.path(getwd(), "unmarked_matrix_360_all.txt"), skip = 1)

#Fit the model 
fit_deer_secr_360_all<-secr.fit(deer_secr_360_all, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_360_all)

save(fit_deer_secr_360_all, file = file.path(getwd(), "fit_deer_secr_360_all.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_360_all<-summary(fit_deer_secr_360_all)[["predicted"]][["estimate"]][[1]]*100
ucl_360_all<-summary(fit_deer_secr_360_all)[["predicted"]][["ucl"]][[1]]*100
lcl_360_all<-summary(fit_deer_secr_360_all)[["predicted"]][["lcl"]][[1]]*100
mean_360_all
ucl_360_all
lcl_360_all

###########################################################
#And now an estimate with just the females
deer_secr_360_females<-read.capthist(captfile = file.path(getwd(), "df_independent_360_females.txt"), 
              trapfile = file.path(getwd(), "camop_360.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_360_females)
plot(deer_secr_360_females, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_360_females<-addSightings(deer_secr_360_females, unmarked = file.path(getwd(), "unmarked_matrix_360_females.txt"), skip = 1)

#Fit the model 
fit_deer_secr_360_females<-secr.fit(deer_secr_360_females, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_360_females)

save(fit_deer_secr_360_females, file = file.path(getwd(), "fit_deer_secr_360_females.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_360_females<-summary(fit_deer_secr_360_females)[["predicted"]][["estimate"]][[1]]*100
ucl_360_females<-summary(fit_deer_secr_360_females)[["predicted"]][["ucl"]][[1]]*100
lcl_360_females<-summary(fit_deer_secr_360_females)[["predicted"]][["lcl"]][[1]]*100
mean_360_females
ucl_360_females
lcl_360_females


```


#################################################################################################
######### POPULATION SIZE: 300 (25% REDUCTION)###################################################
#################################################################################################

Need to re-do the camera operability stuff here because technically its a different run
```{r}
#read in the camop from the 200 run
camop<-read.csv(file.path(getwd(), "camera_operability_300.csv"), header = TRUE)

#This should be really easy since I am programming in perfect operability amongst all cameras. I have the model spit out a file of camera locations, and the number of ticks each camera was operating until

patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#And then transform the camop file to match the format we need:

camop_traps <- camop %>%
  mutate(
    trap.id = camera_id,
    
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    
    # Number of days
    n_days = ticks / (60 * 24),
    
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)

head(camop_traps)

camop_traps<-camop_traps%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```


Ok lets pull out some proportion of individuals and make them marked vs unmarked. Need to introduce some error ya know
```{r}
#read in the lower population size data
df<-read.csv("deer_detections_300.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(75)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```

Export those files to text (needed for secr)
```{r}
write.table(camop_traps, "camop_300.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent, "df_independent_300_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_300_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

########################################################################################
OKAY NOW THE SAME THING AS ABOVE BUT WE ARE ONLY RETAINING INFORMATION FOR FEMALE DEER THROUGHOUT
```{r}
#Read the original frame back in for safety
df<-read.csv("deer_detections_300.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(75)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female") #redundant but I like it

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```


Export those files to text (needed for secr)
```{r}
#Camop file is the same so no need to redo here

write.table(df_independent, "df_independent_300_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_300_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

Now lets fit models for both!
```{r}
#First the full model with all individuals

deer_secr_300_all<-read.capthist(captfile = file.path(getwd(), "df_independent_300_all.txt"), 
              trapfile = file.path(getwd(), "camop_300.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_300_all)
plot(deer_secr_300_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_300_all<-addSightings(deer_secr_300_all, unmarked = file.path(getwd(), "unmarked_matrix_300_all.txt"), skip = 1)

#Fit the model 
fit_deer_secr_300_all<-secr.fit(deer_secr_300_all, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_300_all)

save(fit_deer_secr_300_all, file = file.path(getwd(), "fit_deer_secr_300_all.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_300_all<-summary(fit_deer_secr_300_all)[["predicted"]][["estimate"]][[1]]*100
ucl_300_all<-summary(fit_deer_secr_300_all)[["predicted"]][["ucl"]][[1]]*100
lcl_300_all<-summary(fit_deer_secr_300_all)[["predicted"]][["lcl"]][[1]]*100
mean_300_all
ucl_300_all
lcl_300_all

###########################################################
#And now an estimate with just the females
deer_secr_300_females<-read.capthist(captfile = file.path(getwd(), "df_independent_300_females.txt"), 
              trapfile = file.path(getwd(), "camop_300.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_300_females)
plot(deer_secr_300_females, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_300_females<-addSightings(deer_secr_300_females, unmarked = file.path(getwd(), "unmarked_matrix_300_females.txt"), skip = 1)

#Fit the model 
fit_deer_secr_300_females<-secr.fit(deer_secr_300_females, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_300_females)

save(fit_deer_secr_300_females, file = file.path(getwd(), "fit_deer_secr_300_females.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_300_females<-summary(fit_deer_secr_300_females)[["predicted"]][["estimate"]][[1]]*100
ucl_300_females<-summary(fit_deer_secr_300_females)[["predicted"]][["ucl"]][[1]]*100
lcl_300_females<-summary(fit_deer_secr_300_females)[["predicted"]][["lcl"]][[1]]*100
mean_300_females
ucl_300_females
lcl_300_females


```


#################################################################################################
######### POPULATION SIZE: 200 (50% REDUCTION)###################################################
#################################################################################################

Need to re-do the camera operability stuff here because technically its a different run
```{r}
#read in the camop from the 200 run
camop<-read.csv(file.path(getwd(), "camera_operability_200.csv"), header = TRUE)

#This should be really easy since I am programming in perfect operability amongst all cameras. I have the model spit out a file of camera locations, and the number of ticks each camera was operating until

patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#And then transform the camop file to match the format we need:

camop_traps <- camop %>%
  mutate(
    trap.id = camera_id,
    
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    
    # Number of days
    n_days = ticks / (60 * 24),
    
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)

head(camop_traps)

camop_traps<-camop_traps%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```


Ok lets pull out some proportion of individuals and make them marked vs unmarked. Need to introduce some error ya know
```{r}
#read in the lower population size data
df<-read.csv("deer_detections_200.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
set.seed(420)
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(50)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```

Export those files to text (needed for secr)
```{r}
write.table(camop_traps, "camop_200.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent, "df_independent_200_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_200_all.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

########################################################################################
OKAY NOW THE SAME THING AS ABOVE BUT WE ARE ONLY RETAINING INFORMATION FOR FEMALE DEER THROUGHOUT
```{r}
#Read the original frame back in for safety
df<-read.csv("deer_detections_200.csv", header = TRUE)

#We are only marking females, and only 100 females. Everyone else is unmarked
# 1. Get a list of 100 random unique female IDs
set.seed(420)
marked_female_ids <- df %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(50)

# 2. Create the 'status' column based on that list
df <- df %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids, "marked", "unmarked"))


#Now we create two dataframes: a marked and unmarked. 
df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
#And the marked dataframe, we will treat this as normal below
df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female") #redundant but I like it

######################################################
#But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop_traps$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)



```


Next, format the file of marked detections
```{r}

ticks_per_day <- 60 * 24       # 1440

#My own version
df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)


```


Export those files to text (needed for secr)
```{r}
#Camop file is the same so no need to redo here

write.table(df_independent, "df_independent_200_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix, "unmarked_matrix_200_females.txt", sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```

Now lets fit models for both!
```{r}
#First the full model with all individuals

deer_secr_200_all<-read.capthist(captfile = file.path(getwd(), "df_independent_200_all.txt"), 
              trapfile = file.path(getwd(), "camop_200.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_200_all)
plot(deer_secr_200_all, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_200_all<-addSightings(deer_secr_200_all, unmarked = file.path(getwd(), "unmarked_matrix_200_all.txt"), skip = 1)

#Fit the model 
fit_deer_secr_200_all<-secr.fit(deer_secr_200_all, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_200_all)

save(fit_deer_secr_200_all, file = file.path(getwd(), "fit_deer_secr_200_all.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["estimate"]][[1]]*100
ucl_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["ucl"]][[1]]*100
lcl_200_all<-summary(fit_deer_secr_200_all)[["predicted"]][["lcl"]][[1]]*100
mean_200_all
ucl_200_all
lcl_200_all

###########################################################
#And now an estimate with just the females
deer_secr_200_females<-read.capthist(captfile = file.path(getwd(), "df_independent_200_females.txt"), 
              trapfile = file.path(getwd(), "camop_200.txt"), #A list of files of camera locations for each year
              detector = "count",   #Specify the detector type 
              markocc = rep(0,30), #Specify no marking sessions among the 30 sampling occassions
              skip = 1, 
              verify = TRUE)

   
summary(deer_secr_200_females)
plot(deer_secr_200_females, tracks = TRUE)

#Now add in the unmarked sightings
deer_secr_200_females<-addSightings(deer_secr_200_females, unmarked = file.path(getwd(), "unmarked_matrix_200_females.txt"), skip = 1)

#Fit the model 
fit_deer_secr_200_females<-secr.fit(deer_secr_200_females, 
               buffer = 1000, 
               detectfn = "HN",
               binomN = 0,
               trace = TRUE)

summary(fit_deer_secr_200_females)

save(fit_deer_secr_200_females, file = file.path(getwd(), "fit_deer_secr_200_females.RData"))

#Need to add in the suggested buffer stuff here

#Extract some values
mean_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["estimate"]][[1]]*100
ucl_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["ucl"]][[1]]*100
lcl_200_females<-summary(fit_deer_secr_200_females)[["predicted"]][["lcl"]][[1]]*100
mean_200_females
ucl_200_females
lcl_200_females


```






```{r}
density_figure<-data.frame(mean_density = c(mean_400_all, mean_400_females, mean_360_all, mean_360_females, mean_300_all, mean_300_females, mean_200_all, mean_200_females),
           lc_density = c(lcl_400_all, lcl_400_females, lcl_360_all, lcl_360_females,lcl_300_all, lcl_300_females, lcl_200_all, lcl_200_females),
           uc_density = c(ucl_400_all, ucl_400_females, ucl_360_all, ucl_360_females, ucl_300_all, ucl_300_females, ucl_200_all, ucl_200_females),
           population = factor(c("400", "400", "360", "360", "300", "300", "200", "200"),
                                  levels = c("400", "360", "300", "200")),
           data = c("all", "only females","all", "only females", "all", "only females","all", "only females"))


density_figure<-ggplot(density_figure, aes(x = as.factor(population), 
                           y = mean_density, 
                           color = data, 
                           group = data)) +
  geom_errorbar(aes(ymin = lc_density, ymax = uc_density), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  geom_line(aes(group = data), 
            position = position_dodge(width = 0.5), 
            linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), 
             size = 3) +
  # Add custom labels here:
  scale_x_discrete(labels = c("400" = "Original Population\n(N = 400)", 
                              "360" = "10% Reduction\n(N = 360)", 
                              "300" = "25% Reduction\n(N = 300)", 
                              "200" = "50% Reduction\n(N = 200)")) +
  labs(x = "Simulated Population Size", 
       y = "Estimated Density", 
       color = "Data Subset") +
  theme_minimal()

#Export the figure as well as the main dataframe.
ggsave(filename = "density_figure.jpeg", #name of file
       plot = density_figure,                 #plot you want to save
       path = getwd(),            #where it's saved
       width = 3500,                       #how wide
       height = 2200,                      #how tall
        units = "px",                      #units in pixels
        bg = 'white')                      #make sure background is white


```


```{r}
A<-density_figure%>%
  mutate(population = as.numeric(as.character(A$population)))%>%
  mutate(pop_estimate = mean_density*10)%>%
  #Correct for females (double it)
  mutate(pop_estimate = case_when(data == "only females" ~ pop_estimate*2,
                                   data == "all" ~ pop_estimate))%>%
  mutate(relative_bias = (pop_estimate - population) / population)

A$bias <- A$pop_estimate2 / as.numeric(as.character(A$population))

```