The purpose of this script is to analyze detection data from an agent-based model of deer movement in Oak Bay, British Columbia. This script relies on data from urban_deer_movement_simulation.nlogo, located in the sibling folder 5.1 Netlogo Model. This script produces a series of single session mark resight models in a similar framework to the main paper's analysis. The main outputs of this script are density estimates for each simulated scenario, and a figure comparing density estimates using male & female data vs just female deer data.

Andrew Barnas
andrewbarnas@uvic.ca
April 9th 2026


Load packages and data
```{r}

rm(list=ls())
#Being a bit lazy here, setting working directory to file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

#Packages needed
library(dplyr)        #data wrangling
library(tidyr)        #more data wrangling
library(stringr)      #dealing with strings
library(secr)         #density estimates
library(reproducible) #reproducible workflows
library(ggplot2)      #data visualization

#Setup working directories
netlogo_directory<-reproducible::checkPath(file.path(getwd(), "5.1 Netlogo Model/outputs"), create = TRUE)
input_directory<-reproducible::checkPath(file.path(getwd(), "inputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "outputs"), create = TRUE)
figure_directory<-reproducible::checkPath(file.path(getwd(), "figures"), create = TRUE)


#Read in the datafiles we will need.
#Sepearate detection histories
df_400<-read.csv(file.path(netlogo_directory, "deer_detections_400.csv"), header = TRUE)
df_360<-read.csv(file.path(netlogo_directory, "deer_detections_360.csv"), header = TRUE)
df_300<-read.csv(file.path(netlogo_directory, "deer_detections_300.csv"), header = TRUE)
df_200<-read.csv(file.path(netlogo_directory, "deer_detections_200.csv"), header = TRUE)

#Camera operability
camop_400<-read.csv(file.path(netlogo_directory, "camera_operability_400.csv"), header = TRUE)
camop_360<-read.csv(file.path(netlogo_directory, "camera_operability_360.csv"), header = TRUE)
camop_300<-read.csv(file.path(netlogo_directory, "camera_operability_300.csv"), header = TRUE)
camop_200<-read.csv(file.path(netlogo_directory, "camera_operability_200.csv"), header = TRUE)


```

Ok so in the following sections I do a fair bit of copying and pasting, because 1) I am trying to meet a review deadline, 2) this is easier for now. If I had time I would fancify this, but do as I say not as I do.
But, there are some things I can accomplish once and use them throughout while keeping my environment clean
```{r}
#Patch size of netlogo simulations, used for space calculations
patch_size_m <- 11.18 #roughly
#Rough UTMs based on Oak Bay Zone 10N
utm_x_origin <- 476000
utm_y_origin <- 5362000


#Quickly create the camop_traps file for all scenarios (400, 360, 300, 200)
camop_400 <- camop_400 %>%
  mutate(
    trap.id = camera_id,
    # Convert patch coords -> meters -> UTM
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    # Number of days
    n_days = ticks / (60 * 24),
    # All cameras active every day
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

camop_360 <- camop_360 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))


camop_300 <- camop_300 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))


camop_200 <- camop_200 %>%
  mutate(
    trap.id = camera_id,
    x = utm_x_origin + (xcor * patch_size_m),
    y = utm_y_origin + (ycor * patch_size_m),
    n_days = ticks / (60 * 24),
    usage = str_dup("1", n_days)
  ) %>%
  select(trap.id, x, y, usage)%>%
  arrange(trap.id)%>%
  mutate(x = round(x, digits = 0),
         y = round(y, digits = 0))

```

Another thing I can do seperately to make sure it is version controlled is to isolate the marked females to use for each scenario
```{r}
set.seed(420)
#400 individuals, 200 females, half = 100
marked_female_ids_400 <- df_400 %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(100)
#360 individuals, 180 females, half = 90
marked_female_ids_360 <- df_360 %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(90)
#300 individuals, 150 females, half = 75
marked_female_ids_300 <- df_300 %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(75)
#200 individuals, 100 females, half = 50
marked_female_ids_200 <- df_200 %>%
  filter(sex == "female") %>%
  distinct(deer_id) %>%
  pull(deer_id) %>%
  sample(50)


#And assign the dataframes their marked status
df_400 <- df_400 %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids_400, "marked", "unmarked"))
df_360 <- df_360 %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids_360, "marked", "unmarked"))
df_300 <- df_300 %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids_300, "marked", "unmarked"))
df_200 <- df_200 %>%
  mutate(status = ifelse(deer_id %in% marked_female_ids_200, "marked", "unmarked"))

```




##############################################################################################
Everything past here I am going to deal with seperately our of laziness/ease. For each of the scenarios I need to 1) identify indepenent events for marked and unmarked individuals, 2) format those files for secr, 3) write those files to text

#################################################################################################
######### POPULATION SIZE: 400 (NO REDUCTION)###################################################
#################################################################################################
```{r}


dataframes<-list(df_400, df_360, df_300, df_200)
camops<-list(camop_400, camop_360, camop_300, camop_200)

#A list to store the unmarked matrix counts
unmarked_matrix_list<-list()
#A list to store counts of known individuals
known_individuals_list<-list()

for (j in 1:length(dataframes)) {
  
   # Perform your transformation (placeholder for your actual logic)
  # Let's say you're converting them to matrices
  df <- dataframes[[j]] 
  camop <- camops[[j]]
  
  
  df_unmarked<-df%>%
  filter(status == "unmarked")
 #And the marked dataframe, we will deal with this below
  df_marked<-df%>%
  filter(status == "marked" )
  
  #But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)
 
  
  # Store the result in the i-th position of the list
  unmarked_matrix_list[[j]] <- unmarked_matrix
  
  
  ####################################################
  #Now the marked individuals
  
 df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)
 
 # Store the result in the i-th position of the list
  known_individuals_list[[j]] <- df_independent 
  
}
#Extract unmarked counts
unmarked_matrix_400<-unmarked_matrix_list[[1]]
unmarked_matrix_360<-unmarked_matrix_list[[2]]
unmarked_matrix_300<-unmarked_matrix_list[[3]]
unmarked_matrix_200<-unmarked_matrix_list[[4]]

#extract known counts
df_independent_400<-known_individuals_list[[1]]
df_independent_360<-known_individuals_list[[2]]
df_independent_300<-known_individuals_list[[3]]
df_independent_200<-known_individuals_list[[4]]



```



Now just females!!
#################################################################################################
######### FEMALES ONLY!!!########################################################################
#################################################################################################
```{r}


dataframes<-list(df_400, df_360, df_300, df_200)
camops<-list(camop_400, camop_360, camop_300, camop_200)

#A list to store the unmarked matrix counts
unmarked_matrix_females_list<-list()
#A list to store counts of known individuals
known_individuals_females_list<-list()

for (j in 1:length(dataframes)) {
  
   # Perform your transformation (placeholder for your actual logic)
  # Let's say you're converting them to matrices
  df <- dataframes[[j]] 
  camop <- camops[[j]]
  
  #KEY DIFFERENCE HERE! ONLY RETAINING FEMALE DATA
  df_unmarked<-df%>%
  filter(status == "unmarked")%>%
  filter(sex == "female")
 #And the marked dataframe, we will deal with this below
  df_marked<-df%>%
  filter(status == "marked" )%>%
  filter(sex == "female")
  
  #But carrying on with the unmarked individuals
#First identify "independant detections"
df_unmarked<-df_unmarked%>%
  dplyr::select(-c(deer_id, sex, number.deer, status))%>%
  arrange(camera_id, tick)%>%
  group_by(camera_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))

df_unmarked$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_unmarked),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_unmarked)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_unmarked$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_unmarked$duration[i]) | abs(df_unmarked$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_unmarked$duration[nrow(df_unmarked)] < independence_threshold| 
   is.na(df_unmarked$duration[nrow(df_unmarked)])){
  df_unmarked$Event.ID[nrow(df_unmarked)] <- df_unmarked$Event.ID[nrow(df_unmarked)-1] 
} else{
  df_unmarked$Event.ID[nrow(df_unmarked)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_unmarked<-df_unmarked %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

ticks_per_day <- 60 * 24       # 1440

#Format this a bit more
df_unmarked<-df_unmarked%>% 
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id,  occasion)%>%
  relocate(session, occasion, trap.id)

#############################################################
#Formatting into matrix
# This counts how many times each trap saw an unmarked individual in each window
unmarked_counts <- df_unmarked %>%
  group_by(trap.id, occasion) %>%
  summarise(n = n(), .groups = "drop")

# 2. Create a complete grid of all Traps and all Occasions
# This is crucial so that '0' counts are explicitly represented
all_traps <- camop$trap.id
all_occasions <- 1:30

full_grid <- expand.grid(
  trap.id = all_traps, 
  occasion = all_occasions
)

# 3. Join counts to the grid and pivot to matrix format
unmarked_matrix <- full_grid %>%
  left_join(unmarked_counts, by = c("trap.id", "occasion")) %>%
  mutate(n = replace_na(n, 0)) %>%
  arrange(trap.id, occasion) %>%
  pivot_wider(names_from = occasion, values_from = n) %>%
  select(-trap.id)%>%
  mutate(session = "test")%>%
  relocate(session)
 
  
  # Store the result in the i-th position of the list
  unmarked_matrix_females_list[[j]] <- unmarked_matrix
  
  
  ####################################################
  #Now the marked individuals
  
 df_independent<-df_marked%>%
  arrange(camera_id, tick, deer_id)%>%
  group_by(camera_id, deer_id)%>%
  mutate(duration = as.numeric(tick - lag(tick)))%>%
  dplyr::select(camera_id, tick, deer_id, duration)

df_independent$Event.ID <- 9999
independence_threshold <- 30

seq <- as.numeric(paste0(nrow(df_independent),0))
seq <- round(seq,-(nchar(seq)))


for (i in 2:nrow(df_independent)) {
  #Starting at the first row, assign an event ID that is (E+seq)
  df_independent$Event.ID[i-1]  <- paste0("E",format(seq, scientific = F)) 
  #And then if the next row is an NA (indicating a new individual)
  #Or if the next row is more than 30 mins past the previous row
  #Increase sequence by 1 (so that we now move on to a new Event ID)
  if(is.na(df_independent$duration[i]) | abs(df_independent$duration[i]) > independence_threshold){
    seq <- seq + 1 
  }
}

# Update the information for the last row
# group ID  for the last row
if(df_independent$duration[nrow(df_independent)] < independence_threshold| 
   is.na(df_independent$duration[nrow(df_independent)])){
  df_independent$Event.ID[nrow(df_independent)] <- df_independent$Event.ID[nrow(df_independent)-1] 
} else{
  df_independent$Event.ID[nrow(df_independent)] <- paste0("E",format(seq+1, scientific = F)) 
}

#And lastly, we are just going to rip out a single row from each independent event
#(The first row)
df_independent<-df_independent %>%
  group_by(Event.ID) %>%
  filter(row_number()==1)

#And now convert those to occassions (days)
df_independent<-df_independent%>%
  ungroup()%>%
  mutate(
    occasion = ceiling(tick / ticks_per_day),
    trap.id = camera_id,
    session = "test",
  ) %>%
  select(session, trap.id, deer_id, occasion)%>%
  relocate(session, deer_id, occasion, trap.id)
 
 # Store the result in the i-th position of the list
  known_individuals_females_list[[j]] <- df_independent 
  
}
#Extract unmarked counts
unmarked_matrix_400_females<-unmarked_matrix_females_list[[1]]
unmarked_matrix_360_females<-unmarked_matrix_females_list[[2]]
unmarked_matrix_300_females<-unmarked_matrix_females_list[[3]]
unmarked_matrix_200_females<-unmarked_matrix_females_list[[4]]

#extract known counts
df_independent_400_females<-known_individuals_females_list[[1]]
df_independent_360_females<-known_individuals_females_list[[2]]
df_independent_300_females<-known_individuals_females_list[[3]]
df_independent_200_females<-known_individuals_females_list[[4]]



```

Export all those files! Putting these into an "input directory" for secr models later
```{r}
#Camops
write.table(camop_400, file.path(input_directory, "camop_400.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(camop_360, file.path(input_directory, "camop_360.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(camop_300, file.path(input_directory, "camop_300.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(camop_200, file.path(input_directory, "camop_200.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

###############################
#known individuals
write.table(df_independent_400, file.path(input_directory, "df_independent_400.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(df_independent_400_females, file.path(input_directory, "df_independent_400_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent_360, file.path(input_directory, "df_independent_360.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(df_independent_360_females, file.path(input_directory, "df_independent_360_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent_300, file.path(input_directory, "df_independent_300.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(df_independent_300_females, file.path(input_directory, "df_independent_300_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(df_independent_200, file.path(input_directory, "df_independent_200.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(df_independent_200_females, file.path(input_directory, "df_independent_200_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#Unmarked matrixes
write.table(unmarked_matrix_400, file.path(input_directory, "unmarked_matrix_400.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix_400_females, file.path(input_directory, "unmarked_matrix_400_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix_360, file.path(input_directory, "unmarked_matrix_360.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix_360_females, file.path(input_directory, "unmarked_matrix_360_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


write.table(unmarked_matrix_300, file.path(input_directory, "unmarked_matrix_300.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix_300_females, file.path(input_directory, "unmarked_matrix_300_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


write.table(unmarked_matrix_200, file.path(input_directory, "unmarked_matrix_200.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_matrix_200_females, file.path(input_directory, "unmarked_matrix_200_females.txt"), sep = "\t", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs



```



And now we can fit models for all
```{r}
# Setup your variables
pop_sizes <- c("400", "360", "300", "200")
groups <- c("all", "females")

# Outer loop for population size
for (size in pop_sizes) {
  
  # Inner loop for group (All vs Females)
  for (group in groups) {
    
    # Construct file names dynamically
    # Note: Using the naming logic from your list.files()
    suffix <- ifelse(group == "females", "_females", "")
    
    capt_file <- file.path(input_directory, paste0("df_independent_", size, suffix, ".txt"))
    trap_file <- file.path(input_directory, paste0("camop_", size, ".txt"))
    unmarked_file <- file.path(input_directory, paste0("unmarked_matrix_", size, suffix, ".txt"))
    
    # 1. Read capthist
    deer_capthist <- read.capthist(
      captfile = capt_file, 
      trapfile = trap_file,
      detector = "count",
      markocc = rep(0, 30),
      skip = 1,
      verify = TRUE
    )
    
    # 2. Add unmarked sightings
    deer_capthist <- addSightings(
      deer_capthist, 
      unmarked = unmarked_file, 
      skip = 1
    )
    
    # 3. Fit the model
    # I recommend adding a message so you know which model is running!
    message(paste("Fitting model for size:", size, "| group:", group))
    
    fit_model <- secr.fit(
      deer_capthist, 
      buffer = 1000, 
      detectfn = "HN",
      binomN = 0,
      trace = FALSE # Set to FALSE in loops to keep console clean, or TRUE to monitor
    )
    
    # 4. Save the model object
    # We create a unique filename like "fit_deer_secr_400_females.RData"
    out_filename <- paste0("fit_deer_secr_", size, "_", group, ".RData")
    save(fit_model, file = file.path(output_directory, out_filename))
    
    # Clean up memory after each fit
    rm(fit_model, deer_capthist)
    gc()
  }
}




```

Bring the saved model objects back in, extract estimates and plot!
```{r}
pop_sizes <- c("400", "360", "300", "200")
groups <- c("all", "females")

# Create an empty dataframe to store results
results <- data.frame()

for (size in pop_sizes) {
  for (group in groups) {
    
    file_name <- paste0("fit_deer_secr_", size, "_", group, ".RData")
    file_path <- file.path(output_directory, file_name)
    
    # Only proceed if the file actually exists
    if (file.exists(file_path)) {
      
      # Load the file. Remember, inside the loop it was saved as 'fit_model'
      load(file_path)
      
      # Extract predicted density (usually the first row of the 'predicted' table)
      stats <- summary(fit_model)$predicted
      
      # Create a temporary row
      temp_df <- data.frame(
        population_size = as.numeric(size),
        sex_group = group,
        density = stats["D", "estimate"] * 100,
        ucl = stats["D", "ucl"] * 100,
        lcl = stats["D", "lcl"] * 100
      )
      
      # Bind to the main results table
      results <- rbind(results, temp_df)
      
      # Clean up to keep the environment tidy
      rm(fit_model)
    }
  }
}

#And plot
ggplot(results, aes(x = as.factor(population_size), 
                           y = density, 
                           color = sex_group, 
                           group = sex_group)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), 
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  geom_line(aes(group = sex_group), 
            position = position_dodge(width = 0.5), 
            linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), 
             size = 3) +
  # Add custom labels here:
  scale_x_discrete(labels = c("400" = "Original Population\n(N = 400)", 
                              "360" = "10% Reduction\n(N = 360)", 
                              "300" = "25% Reduction\n(N = 300)", 
                              "200" = "50% Reduction\n(N = 200)")) +
  labs(x = "Simulated Population Size", 
       y = "Estimated Density", 
       color = "Data Subset") +
  theme_minimal()

#Export the figure as well as the main dataframe.
# ggsave(filename = "density_figure.jpeg", #name of file
#        plot = density_figure,                 #plot you want to save
#        path = getwd(),            #where it's saved
#        width = 3500,                       #how wide
#        height = 2200,                      #how tall
#         units = "px",                      #units in pixels
#         bg = 'white')                      #make sure background is white




```




