The purpose of this script is to take the outputs from 1.2 Data Prep and format the data for input into the SECR models. This is similar to the data prep step before, but specialized for secr. The main outputs of this for each year are:
1. A capture list of known individuals
2. A camera location and operability file
3. A matrix of counts of unmarked individuals

Also, in addition to producing individual years of data, we also output the same information but in a format for multisession models (amalgamated together)

Andrew Barnas - May 2024
andrew.f.barnas@gmail.com

Load packages and data we need
```{r}
#Clear everything out, its probably messy from the data prep file
rm(list = ls())

#Packages we will need
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(lubridate)    #datetime data
library(ggplot2)      #data visualization
library(stringr)      #reading information on strings
library(secr)         #the actual model engine
library(sf)        #dealing with spatial data

#Set the working directories to read in the files we need
input_directory<-reproducible::checkPath(file.path(getwd(), "1.2 Data Prep/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "2. SECR Prep/outputs"), create = TRUE)

#Detection files
deer_2018<-read.csv(file.path(input_directory,"marked_detections_sept_2018.csv"))
deer_2019<-read.csv(file.path(input_directory,"marked_detections_fall_2019.csv"))
deer_2020<-read.csv(file.path(input_directory,"marked_detections_fall_2020.csv"))
deer_2021<-read.csv(file.path(input_directory,"marked_detections_sept_2021.csv"))
deer_2022<-read.csv(file.path(input_directory,"marked_detections_sept_2022.csv"))
deer_2023<-read.csv(file.path(input_directory,"marked_detections_sept_2023.csv"))

#Matrix of unmarked detection counts
unmarked_2018<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2018.csv"))
unmarked_2019<-read.csv(file.path(input_directory,"unmarked_matrix_fall_2019.csv"))
unmarked_2020<-read.csv(file.path(input_directory,"unmarked_matrix_fall_2020.csv"))
unmarked_2021<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2021.csv"))
unmarked_2022<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2022.csv"))
unmarked_2023<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2023.csv"))

#Camera operation matrix
usage_2018<-read.csv(file.path(input_directory,"operation_matrix_sept_2018.csv"))
usage_2019<-read.csv(file.path(input_directory,"operation_matrix_fall_2019.csv"))
usage_2020<-read.csv(file.path(input_directory,"operation_matrix_fall_2020.csv"))
usage_2021<-read.csv(file.path(input_directory,"operation_matrix_sept_2021.csv"))
usage_2022<-read.csv(file.path(input_directory,"operation_matrix_sept_2022.csv"))
usage_2023<-read.csv(file.path(input_directory,"operation_matrix_sept_2023.csv"))

```

Lets deal with this year by year, so you don't have to run through everything each time.

#######################################################################################
#######################################################################################
#######################################################################################


FORMATTING 2018 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2018<-data.frame(occassion = seq(ymd("2018-09-01"), ymd("2018-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2018<-deer_2018%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2018<- data.frame(SiteID= usage_2018$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2018$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2018<-data.frame(WHID_all = unique(deer_2018$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2018$WHID_all), by = 1))

deer_2018<-merge(deer_2018, sequence_2018, by = "occassion")
deer_2018<-merge(deer_2018, trap_names_2018, by = "SiteID")
deer_2018<-merge(deer_2018, wildlife_ids_2018, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2018<-deer_2018%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2018<-usage_2018%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2018<-merge(cams_2018, trap_names_2018, by = "SiteID")
cams_2018<-cams_2018%>%
  dplyr::select(trap.id = trap.id_number, x, y)

library(sf)
x<-st_as_sf(cams_2018, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2018$trap.id_number)

#Old code using library(rgdal)
#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
#x<-SpatialPoints(data.frame(cams_2018$y, cams_2018$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
#proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
#x<-spTransform(x, CRS = "+proj=utm +zone=10
#+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
#coordinates<-data.frame(x@coords)
#coordinates<-coordinates%>%
 # mutate(trap.id = trap_names_2018$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2018<-merge(cams_2018, coordinates, by = "trap.id")
cams_2018<-cams_2018%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2018
#Just need to collapse across all the occassions 
cams_2018$usage<-apply( usage_2018[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2018$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2018
#Need to clean it up a bit
unmarked_2018<-unmarked_2018%>%
  mutate(session = "Oakbay2018")%>%
  dplyr::select(session, 2:ncol(unmarked_2018))

#Fix the column names
colnames(unmarked_2018)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2019 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2019<-data.frame(occassion = seq(ymd("2019-10-08"), ymd("2019-11-11"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:35))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2019<-deer_2019%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2019<- data.frame(SiteID= usage_2019$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2019$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2019<-data.frame(WHID_all = unique(deer_2019$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2019$WHID_all), by = 1))

deer_2019<-merge(deer_2019, sequence_2019, by = "occassion")
deer_2019<-merge(deer_2019, trap_names_2019, by = "SiteID")
deer_2019<-merge(deer_2019, wildlife_ids_2019, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2019<-deer_2019%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2019<-usage_2019%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2019<-merge(cams_2019, trap_names_2019, by = "SiteID")
cams_2019<-cams_2019%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#library(sf)
x<-st_as_sf(cams_2019, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2019$trap.id_number)

#Old code using library(rgdal)
#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
#x<-SpatialPoints(data.frame(cams_2018$y, cams_2018$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
#proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
#x<-spTransform(x, CRS = "+proj=utm +zone=10
#+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
#coordinates<-data.frame(x@coords)
#coordinates<-coordinates%>%
 # mutate(trap.id = trap_names_2018$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2019<-merge(cams_2019, coordinates, by = "trap.id")
cams_2019<-cams_2019%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2019
#Just need to collapse across all the occassions 
cams_2019$usage<-apply( usage_2019[, 4:38], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2019$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2019
unmarked_2019
#Need to clean it up a bit
unmarked_2019<-unmarked_2019%>%
  mutate(session = "Oakbay2019")%>%
  dplyr::select(session, 2:ncol(unmarked_2019))

#Fix the column names
colnames(unmarked_2019)<-c("session", 1:35)

```
#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2020 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2020<-data.frame(occassion = seq(ymd("2020-10-08"), ymd("2020-11-11"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:35))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2020<-deer_2020%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2020<- data.frame(SiteID= usage_2020$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2020$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2020<-data.frame(WHID_all = unique(deer_2020$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2020$WHID_all), by = 1))

deer_2020<-merge(deer_2020, sequence_2020, by = "occassion")
deer_2020<-merge(deer_2020, trap_names_2020, by = "SiteID")
deer_2020<-merge(deer_2020, wildlife_ids_2020, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2020<-deer_2020%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2020<-usage_2020%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2020<-merge(cams_2020, trap_names_2020, by = "SiteID")
cams_2020<-cams_2020%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#library(sf)
x<-st_as_sf(cams_2020, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2020$trap.id_number)

#Old code using library(rgdal)
#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
#x<-SpatialPoints(data.frame(cams_2018$y, cams_2018$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
#proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
#x<-spTransform(x, CRS = "+proj=utm +zone=10
#+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
#coordinates<-data.frame(x@coords)
#coordinates<-coordinates%>%
 # mutate(trap.id = trap_names_2018$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2020<-merge(cams_2020, coordinates, by = "trap.id")
cams_2020<-cams_2020%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2020
#Just need to collapse across all the occassions 
cams_2020$usage<-apply( usage_2020[, 4:38], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2020$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2020
unmarked_2020
#Need to clean it up a bit
unmarked_2020<-unmarked_2020%>%
  mutate(session = "Oakbay2020")%>%
  dplyr::select(session, 2:ncol(unmarked_2020))

#Fix the column names
colnames(unmarked_2020)<-c("session", 1:35)

```
#######################################################################################
#######################################################################################
#######################################################################################


FORMATTING 2021 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2021<-data.frame(occassion = seq(ymd("2021-09-01"), ymd("2021-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2021<-deer_2021%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2021<- data.frame(SiteID= usage_2021$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2021$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2021<-data.frame(WHID_all = unique(deer_2021$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2021$WHID_all), by = 1))

deer_2021<-merge(deer_2021, sequence_2021, by = "occassion")
deer_2021<-merge(deer_2021, trap_names_2021, by = "SiteID")
deer_2021<-merge(deer_2021, wildlife_ids_2021, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2021<-deer_2021%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2021<-usage_2021%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2021<-merge(cams_2021, trap_names_2021, by = "SiteID")
cams_2021<-cams_2021%>%
  dplyr::select(trap.id = trap.id_number, x, y)

x<-st_as_sf(cams_2021, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2021$trap.id_number)

#And merge the new coordinates back into the cam dataframe
cams_2021<-merge(cams_2021, coordinates, by = "trap.id")
cams_2021<-cams_2021%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2021
#Just need to collapse across all the occassions 
cams_2021$usage<-apply( usage_2021[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2021$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2021
#Need to clean it up a bit
unmarked_2021<-unmarked_2021%>%
  mutate(session = "Oakbay2021")%>%
  dplyr::select(session, 2:ncol(unmarked_2021))

#Fix the column names
colnames(unmarked_2021)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################


FORMATTING 2022 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2022<-data.frame(occassion = seq(ymd("2022-09-01"), ymd("2022-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2022<-deer_2022%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2022<- data.frame(SiteID= usage_2022$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2022$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2022<-data.frame(WHID_all = unique(deer_2022$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2022$WHID_all), by = 1))

deer_2022<-merge(deer_2022, sequence_2022, by = "occassion")
deer_2022<-merge(deer_2022, trap_names_2022, by = "SiteID")
deer_2022<-merge(deer_2022, wildlife_ids_2022, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2022<-deer_2022%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2022<-usage_2022%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2022<-merge(cams_2022, trap_names_2022, by = "SiteID")
cams_2022<-cams_2022%>%
  dplyr::select(trap.id = trap.id_number, x, y)

x<-st_as_sf(cams_2022, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2022$trap.id_number)

#And merge the new coordinates back into the cam dataframe
cams_2022<-merge(cams_2022, coordinates, by = "trap.id")
cams_2022<-cams_2022%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2022
#Just need to collapse across all the occassions 
cams_2022$usage<-apply( usage_2022[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2022$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2022
#Need to clean it up a bit
unmarked_2022<-unmarked_2022%>%
  mutate(session = "Oakbay2022")%>%
  dplyr::select(session, 2:ncol(unmarked_2022))

#Fix the column names
colnames(unmarked_2022)<-c("session", 1:30)

```


#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2023 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2023<-data.frame(occassion = seq(ymd("2023-09-01"), ymd("2023-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2023<-deer_2023%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2023<- data.frame(SiteID= usage_2023$SiteID,
                        trap.id_number = seq(from = 1, to = n_distinct(usage_2023$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2023<-data.frame(WHID_all = unique(deer_2023$WHID_all),
                         id_number = seq(from = 1, to = n_distinct(deer_2023$WHID_all), by = 1))

deer_2023<-merge(deer_2023, sequence_2023, by = "occassion")
deer_2023<-merge(deer_2023, trap_names_2023, by = "SiteID")
deer_2023<-merge(deer_2023, wildlife_ids_2023, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2023<-deer_2023%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2023<-usage_2023%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2023<-merge(cams_2023, trap_names_2023, by = "SiteID")
cams_2023<-cams_2023%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#secr demands coordinates are projected to the xy plane, so we need to do a couple things
x<-st_as_sf(cams_2023, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = trap_names_2023$trap.id_number)

#And merge the new coordinates back into the cam dataframe
cams_2023<-merge(cams_2023, coordinates, by = "trap.id")
cams_2023<-cams_2023%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2023
#Just need to collapse across all the occassions 
cams_2023$usage<-apply( usage_2023[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2023$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2023
#Need to clean it up a bit
unmarked_2023<-unmarked_2023%>%
  mutate(session = "Oakbay2023")%>%
  dplyr::select(session, 2:ncol(unmarked_2023))

#Fix the column names
colnames(unmarked_2023)<-c("session", 1:30)

```

#######################################################################################
#######################################################################################
#######################################################################################


OUTPUTTING TO TEXT FILES!!!
```{r}
#OUTPUTTING TO TEXT FILES
#secr demands that files are read in as text files, so going to output these to our output directory

########################################################################################################
#2018 Files
#Trap file - camera locations and usage
write.table(cams_2018, file.path(output_directory, "deer_cameras_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(deer_2018, file.path(output_directory, "captures_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(unmarked_2018, file.path(output_directory, "unmarked_matrix_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


########################################################################################################
#2019 Files
#Trap file - camera locations and usage
write.table(cams_2019, file.path(output_directory, "deer_cameras_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(deer_2019, file.path(output_directory, "captures_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(unmarked_2019, file.path(output_directory, "unmarked_matrix_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

########################################################################################################
#2020 Files
#Trap file - camera locations and usage
write.table(cams_2020, file.path(output_directory, "deer_cameras_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(deer_2020, file.path(output_directory, "captures_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(unmarked_2020, file.path(output_directory, "unmarked_matrix_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs




########################################################################################################
#2021 Files
#Trap file - camera locations and usage
write.table(cams_2021, file.path(output_directory, "deer_cameras_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2021, file.path(output_directory, "captures_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2021, file.path(output_directory, "unmarked_matrix_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

########################################################################################################
#2022 Files
#Trap file - camera locations and usage
write.table(cams_2022, file.path(output_directory, "deer_cameras_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2022, file.path(output_directory, "captures_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2022, file.path(output_directory, "unmarked_matrix_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


########################################################################################################
#2023 Files
#Trap file - camera locations and usage
write.table(cams_2023, file.path(output_directory, "deer_cameras_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2023, file.path(output_directory, "captures_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2023, file.path(output_directory, "unmarked_matrix_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#MULTI-SESSION OUTPUT
# write.table(deer_all, file.path(output_directory, "deer_multisession.txt"), sep = " ", 
#             quote = FALSE,     #no quotes allowed!
#             row.names = FALSE) #prevent including row IDs
# 
# write.table(deer_reduced, file.path(output_directory, "deer_reduced.txt"), sep = " ", 
#             quote = FALSE,     #no quotes allowed!
#             row.names = FALSE) #prevent including row IDs
# 
# 
# #Quickly whip up and unmarked matrix
# unmarked_matrix<-rbind(unmarked_2021, unmarked_2022)
# write.table(unmarked_matrix, file.path(output_directory, "unmarked_matrix.txt"), sep = " ", 
#             quote = FALSE,     #no quotes allowed!
#             row.names = FALSE) #prevent including row IDs
```

EVERYTHING PAST THIS POINT IS OLD AND WILL NEED TO BE UPDATED WITH THE NEW DATA
################################################################################

















#############################
2020 DATA ###################
#############################
FORMATTING TRAP FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2020<-data.frame(occassion = seq(ymd("2020-09-01"), ymd("2020-09-30"), by = "day"), #These are the days included in the file
                          occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2020<-deer_2020%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2020<- data.frame(SiteID= usage_2020$SiteID,
                             trap.id_number = seq(from = 1, to = n_distinct(usage_2020$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2020<-data.frame(WHID_all = unique(deer_2020$WHID_all),
                              id_number = seq(from = 1, to = n_distinct(deer_2020$WHID_all), by = 1))

deer_2020<-merge(deer_2020, sequence_2020, by = "occassion")
deer_2020<-merge(deer_2020, trap_names_2020, by = "SiteID")
deer_2020<-merge(deer_2020, wildlife_ids_2020, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2020<-deer_2020%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)

```

FORMATTING TRAP FILES
```{r}
#Clean up the location file
cams_2020<-usage_2020%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2020<-merge(cams_2020, trap_names_2020, by = "SiteID")
cams_2020<-cams_2020%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
x<-SpatialPoints(data.frame(cams_2020$y, cams_2020$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
x<-spTransform(x, CRS = "+proj=utm +zone=10
+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
coordinates<-data.frame(x@coords)
coordinates<-coordinates%>%
  mutate(trap.id = trap_names_2020$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2020<-merge(cams_2020, coordinates, by = "trap.id")
cams_2020<-cams_2020%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = cams_2020.y, y=cams_2020.x) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2020
#Just need to collapse across all the occassions 
cams_2020$usage<-apply( usage_2020[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2020$usage)

```

UNMARKED MATRIX 
```{r}
#TU - unmarked matrix 2021
unmarked_2020
#Need to clean it up a bit
unmarked_2020<-unmarked_2020%>%
  mutate(session = "Oakbay2020")%>%
  dplyr::select(session, 2:ncol(unmarked_2020))

#Fix the column names
colnames(unmarked_2020)<-c("session", 1:30)

```


#############################
2021 DATA ###################
#############################


```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2021<-data.frame(occassion = seq(ymd("2021-09-01"), ymd("2021-09-30"), by = "day"), #These are the days included in the file
                          occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2021<-deer_2021%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2021<- data.frame(SiteID= usage_2021$SiteID,
                             trap.id_number = seq(from = 1, to = n_distinct(usage_2021$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2021<-data.frame(WHID_all = unique(deer_2021$WHID_all),
                              id_number = seq(from = 1, to = n_distinct(deer_2021$WHID_all), by = 1))

deer_2021<-merge(deer_2021, sequence_2021, by = "occassion")
deer_2021<-merge(deer_2021, trap_names_2021, by = "SiteID")
deer_2021<-merge(deer_2021, wildlife_ids_2021, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2021<-deer_2021%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)

```

FORMATTING TRAP FILES
```{r}
#Clean up the location file
cams_2021<-usage_2021%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2021<-merge(cams_2021, trap_names_2021, by = "SiteID")
cams_2021<-cams_2021%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
x<-SpatialPoints(data.frame(cams_2021$y, cams_2021$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
x<-spTransform(x, CRS = "+proj=utm +zone=10
+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
coordinates<-data.frame(x@coords)
coordinates<-coordinates%>%
  mutate(trap.id = trap_names_2021$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2021<-merge(cams_2021, coordinates, by = "trap.id")
cams_2021<-cams_2021%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = cams_2021.y, y=cams_2021.x) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2021
#Just need to collapse across all the occassions 
cams_2021$usage<-apply( usage_2021[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2021$usage)

```

UNMARKED MATRIX 
```{r}
#TU - unmarked matrix 2021
unmarked_2021
#Need to clean it up a bit
unmarked_2021<-unmarked_2021%>%
  mutate(session = "Oakbay2021")%>%
  dplyr::select(session, 2:ncol(unmarked_2021))

#Fix the column names
colnames(unmarked_2021)<-c("session", 1:30)

```

#############################
2022 DATA ###################
#############################
FORMATTING TRAP FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2022<-data.frame(occassion = seq(ymd("2022-09-01"), ymd("2022-09-30"), by = "day"), #These are the days included in the file
                          occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2022<-deer_2022%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
trap_names_2022<- data.frame(SiteID= usage_2022$SiteID,
                             trap.id_number = seq(from = 1, to = n_distinct(usage_2022$SiteID), by = 1))

#Create a frame of wildlife ID numbers
wildlife_ids_2022<-data.frame(WHID_all = unique(deer_2022$WHID_all),
                              id_number = seq(from = 1, to = n_distinct(deer_2022$WHID_all), by = 1))

deer_2022<-merge(deer_2022, sequence_2022, by = "occassion")
deer_2022<-merge(deer_2022, trap_names_2022, by = "SiteID")
deer_2022<-merge(deer_2022, wildlife_ids_2022, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2022<-deer_2022%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap.id_number)

```

FORMATTING TRAP FILES
```{r}
#Clean up the location file
cams_2022<-usage_2022%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2022<-merge(cams_2022, trap_names_2022, by = "SiteID")
cams_2022<-cams_2022%>%
  dplyr::select(trap.id = trap.id_number, x, y)

#secr demands coordinates are projected to the xy plane, so we need to do a couple things
#First we have to read the decimal degrees format in as a spatial object with SpatialPoints
x<-SpatialPoints(data.frame(cams_2022$y, cams_2022$x)) #I cant get this to work in x y format...
#Assign a coordinate system (going with good old WGS84)
proj4string(x)<- "+proj=longlat +datum=WGS84 +no_defs + ellps=WGS84 +towgs84=0,0,0"
#And now transform those coordinates to UTM (a projected coordinate system)
#Transform it
x<-spTransform(x, CRS = "+proj=utm +zone=10
+ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0")
#Extract the relevant data
coordinates<-data.frame(x@coords)
coordinates<-coordinates%>%
  mutate(trap.id = trap_names_2022$trap.id_number) #Since they are in order I think you can do this

#And merge the new coordinates back into the cam dataframe
cams_2022<-merge(cams_2022, coordinates, by = "trap.id")
cams_2022<-cams_2022%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = cams_2022.y, y=cams_2022.x) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2022
#Just need to collapse across all the occassions 
cams_2022$usage<-apply( usage_2022[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2022$usage)

```

UNMARKED MATRIX 
```{r}
#TU - unmarked matrix 2021
unmarked_2022
#Need to clean it up a bit
unmarked_2022<-unmarked_2022%>%
  mutate(session = "Oakbay2022")%>%
  dplyr::select(session, 2:ncol(unmarked_2022))

#Fix the column names
colnames(unmarked_2022)<-c("session", 1:30)

```

Multi-session data prep
```{r}
#Here I want to take the individual session data and put them into a single file
#Reminding me which files I need to merge
deer_2018
deer_2020
deer_2021
deer_2022

#deer_all<-rbind(deer_2018, deer_2020, deer_2021, deer_2022)

deer_reduced<-rbind(deer_2021, deer_2022)

#What about traps?

```



```{r}
#OUTPUTTING TO TEXT FILES
#secr demands that files are read in as text files, so going to output these to our output directory

#2018 Files
#Trap file - camera locations and usage
write.table(cams_2018, file.path(output_directory, "deer_cameras_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(deer_2018, file.path(output_directory, "captures_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(unmarked_2018, file.path(output_directory, "unmarked_matrix_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#2019 Files
#Trap file - camera locations and usage
write.table(cams_2019, file.path(output_directory, "deer_cameras_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(deer_2019, file.path(output_directory, "captures_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
#Marked capture file - detections of known individuals 
write.table(unmarked_2019, file.path(output_directory, "unmarked_matrix_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#2020 Files
#Trap file - camera locations and usage
write.table(cams_2020, file.path(output_directory, "deer_cameras_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2020, file.path(output_directory, "captures_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2020, file.path(output_directory, "unmarked_matrix_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#2021 Files
#Trap file - camera locations and usage
write.table(cams_2021, file.path(output_directory, "deer_cameras_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2021, file.path(output_directory, "captures_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2021, file.path(output_directory, "unmarked_matrix_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#2022 Files
#Trap file - camera locations and usage
write.table(cams_2022, file.path(output_directory, "deer_cameras_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(deer_2022, file.path(output_directory, "captures_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Marked capture file - detections of known individuals 
write.table(unmarked_2022, file.path(output_directory, "unmarked_matrix_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#MULTI-SESSION OUTPUT
write.table(deer_all, file.path(output_directory, "deer_multisession.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_reduced, file.path(output_directory, "deer_reduced.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#Quickly whip up and unmarked matrix
unmarked_matrix<-rbind(unmarked_2021, unmarked_2022)
write.table(unmarked_matrix, file.path(output_directory, "unmarked_matrix.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
```