TEST DATA PREP FOR MULTI SESSION 

TESTING WITH 2020 AND 2021 BECAUSE THEY HAVE DIFFERING NUMBER OF DAYS AND I NEED TO FIGURE OUT HOW TO RECONCILE

Load packages and data we need
```{r}
#Clear everything out, its probably messy from the data prep file
rm(list = ls())

#Packages we will need
library(reproducible) #setting up directories for data management
library(dplyr)        #data wranglin'
library(tidyr)
library(lubridate)    #datetime data
library(ggplot2)      #data visualization
library(stringr)      #reading information on strings
library(secr)         #the actual model engine
library(sf)        #dealing with spatial data

#Set the working directories to read in the files we need
input_directory<-reproducible::checkPath(file.path(getwd(), "1.2 Data Prep/outputs"), create = TRUE)
output_directory<-reproducible::checkPath(file.path(getwd(), "2. SECR Prep/outputs/multisession"), create = TRUE)
mortality_data<-reproducible::checkPath(file.path(getwd(), "1.1 Assigning WHIDs/inputs"), create = TRUE)

#Detection files
deer_2018<-read.csv(file.path(input_directory,"marked_detections_sept_2018.csv"))
deer_2019<-read.csv(file.path(input_directory,"marked_detections_fall_2019.csv"))
deer_2020<-read.csv(file.path(input_directory,"marked_detections_fall_2020.csv"))
deer_2021<-read.csv(file.path(input_directory,"marked_detections_sept_2021.csv"))
deer_2022<-read.csv(file.path(input_directory,"marked_detections_sept_2022.csv"))
deer_2023<-read.csv(file.path(input_directory,"marked_detections_sept_2023.csv"))

#Matrix of unmarked detection counts
unmarked_2018<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2018.csv"))
unmarked_2019<-read.csv(file.path(input_directory,"unmarked_matrix_fall_2019.csv"))
unmarked_2020<-read.csv(file.path(input_directory,"unmarked_matrix_fall_2020.csv"))
unmarked_2021<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2021.csv"))
unmarked_2022<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2022.csv"))
unmarked_2023<-read.csv(file.path(input_directory,"unmarked_matrix_sept_2023.csv"))

#Camera operation matrix
usage_2018<-read.csv(file.path(input_directory,"operation_matrix_sept_2018.csv"))
usage_2019<-read.csv(file.path(input_directory,"operation_matrix_fall_2019.csv"))
usage_2020<-read.csv(file.path(input_directory,"operation_matrix_fall_2020.csv"))
usage_2021<-read.csv(file.path(input_directory,"operation_matrix_sept_2021.csv"))
usage_2022<-read.csv(file.path(input_directory,"operation_matrix_sept_2022.csv"))
usage_2023<-read.csv(file.path(input_directory,"operation_matrix_sept_2023.csv"))

#Mortality Data
mortality<-read.csv(file.path(mortality_data, "ubran_deer_mortalities.csv"), header = TRUE)

```

FORMATTING TRAP NAMES FOR ALL
```{r}
#So I think the first thing we need to do is create a giant thing of all the trap names
trap_names <- data.frame(SiteID = (c(unique(usage_2018$SiteID),
                                     unique(usage_2019$SiteID),
                                     unique(usage_2020$SiteID), 
                                     unique(usage_2021$SiteID),
                                     unique(usage_2022$SiteID),
                                     unique(usage_2023$SiteID))))%>%
  group_by(SiteID)%>%
  slice(1)

trap_names<-data.frame(trap_names,
           trap_id_number = seq(from = 1, to = n_distinct(trap_names$SiteID), by = 1))


#And same with Wildlife Health ID numbers - this is everyone in the study
id_names<-data.frame(WHID_all = c(unique(deer_2018$WHID_all),
                                  unique(deer_2019$WHID_all),
                                  unique(deer_2020$WHID_all), 
                                  unique(deer_2021$WHID_all),
                                  unique(deer_2022$WHID_all),
                                  unique(deer_2023$WHID_all)))%>%
  group_by(WHID_all)%>%
  slice(1)

id_names<-data.frame(id_names, 
                     id_number = seq(from = 1, to = n_distinct(id_names$WHID_all)))


```

Prepping the mortality data- There are two things that need to be done here. For every individual in the study, I need to indicate the year they were first available to be detected, and then a cutoff date for when they were no longer available. This will involve making a matrix for each session/year, with a row of data for each ordered individual id, and a column for each occassion (1 to 30). 
```{r}
#First, I can just clean up the mortality data a little
mortality<-mortality%>%
  dplyr::select(WHID_all = WHID,
                mortality_date = Mortality.Date)%>%
  group_by(WHID_all)%>%
  summarise(mortality_date = max(mortality_date))

#No doubles? Good. 
mortality%>%
  group_by(WHID_all)%>%
  summarise(n_obs = n())%>%
  filter(n_obs > 1)


  

#Removing that one observation that has no date, I can't use this.
mortality<-mortality%>%
  mutate(mortality_date = case_when(str_detect(mortality_date, "not dead") ~ "",
                                    TRUE ~ mortality_date))


#One quick thing here, the replacements have capital Rs, but should fix that
mortality<-mortality%>%
  mutate(WHID_all = str_replace(WHID_all, pattern = "R", replacement = "r"))

#First we need to rename the WHID_all category
mortality<-mortality%>%
  mutate(WHID_all = str_c("['", WHID_all, "']"))

#####################################################################
#Ok now that its cleaned
#First, who do we have detection data for that we have not included in the mortality data?
id_names%>%
  filter(!WHID_all %in% unique(mortality$WHID_all))

#Alright diving back into the original files, it looks like these two where originally a different collar so I can rename them
id_names<-id_names%>%
  mutate(WHID_all = case_when(str_detect(WHID_all, pattern = "20-0129") ~ "['18-11402']",
                              str_detect(WHID_all, pattern = "20-0134") ~ "['18-12479']",
                              TRUE ~ WHID_all))

#Are there any mortality names that are not in the id_names? yes obviously, they were not detected.
mortality%>%
  filter(!WHID_all %in% unique(id_names$WHID_all))

#Ok here is what I am going to do, lets quickly make a new id names file. This is based off of everyone.
id_names<-data.frame(mortality, 
                     id_number = seq(from = 1, to = n_distinct(mortality$WHID_all)))%>%
  select(WHID_all, id_number)
#######################################################################################################
#So now we can start making the matrices for each year

#Ok so now what I need is a sort of start date for everyone.
#I guess what I can do is make the start date Jan 1st for each year indicated by the number?
mortality<-mortality%>%
  mutate(first_date = case_when(str_detect(WHID_all, "18-") ~ "01-Jan-2018",
                                str_detect(WHID_all, "19-") ~ "01-Jan-2019",
                                str_detect(WHID_all, "20-") ~ "01-Jan-2020"))%>%
  relocate(WHID_all, first_date, mortality_date)
  #If there is no confirmed death date, we assume they are in the population forever
mortality[mortality ==""]<-"01-Jan-2025"

####################################################################################################
#Ok so now for each year what I need to do is figure out what days of the sampling period these deer were available for detection
#read in the date times
#Need to do this step ahead of time because of that one NA that wont go away
A<-mortality%>%
  mutate(mortality_date = case_when(WHID_all == "['20-0133']" ~ "2025-01-01",
                                    TRUE ~ mortality_date))

#And now we can parse in the dates
A<-A%>%
  mutate(first_date = dmy(first_date),
         mortality_date = ymd(parse_date_time(mortality_date, orders = c("dmy", "ymd"))))

#Lets replace that one that didn't parse - just make it the latest date
#I hate doing this manually but


#Ok, this makes a sequence of dates for the whole study period. We want to use this sequence of dates, alongside the sequence of dates that each deer was available to be sighted, and indicate whether a deer could have been sighted within the sequence of dates
earliest_date<-min(A$first_date)
latest_date<-max(A$mortality_date)
date.sequence<-seq(earliest_date, latest_date, by = "day")

#Now lets create a sequence of dates for each deer
B<-A%>%
  rowwise()%>%
  do(data.frame(WHID_all = .$WHID_all, date = seq(.$first_date, .$mortality_date, by = "day")))%>%
  mutate(available = 1)

#And now we make a dataframe that includes every combination of the possible days, and the days the deer were available. 
date_deer_expanded<-expand.grid(date = date.sequence, WHID_all = unique(A$WHID_all))
#And now merge them, add in zeros for not being available
B<-merge(date_deer_expanded, B, by = c("WHID_all", "date"), all.x = TRUE)%>%
  replace(is.na(.), 0)

#Ok and so now, I think I can just make the matrices for each year
markocc_2018<-B%>%
  filter(date >= ymd("2018-09-01") & date <= ymd("2018-09-30"))%>%
  tidyr::spread(date, value = available)
  #select(-WHID_all)

markocc_2019<-B%>%
  filter(date >= ymd("2019-10-08") & date <= ymd("2019-11-06"))%>%
  tidyr::spread(date, value = available)
  #select(-WHID_all)

markocc_2020<-B%>%
  filter(date >= ymd("2020-10-08") & date <= ymd("2020-11-06"))%>%
  tidyr::spread(date, value = available)
  #select(-WHID_all)

markocc_2021<-B%>%
  filter(date >= ymd("2021-09-01") & date <= ymd("2021-09-30"))%>%
  tidyr::spread(date, value = available)
 # select(-WHID_all)

markocc_2022<-B%>%
  filter(date >= ymd("2022-09-01") & date <= ymd("2022-09-30"))%>%
  tidyr::spread(date, value = available)
 # select(-WHID_all)

markocc_2023<-B%>%
  filter(date >= ymd("2023-09-01") & date <= ymd("2023-09-30"))%>%
  tidyr::spread(date, value = available)
  #select(-WHID_all)



```



#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2018 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2018<-data.frame(occassion = seq(ymd("2018-09-01"), ymd("2018-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2018<-deer_2018%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2018<- data.frame(SiteID= usage_2018$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2018$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2018<-data.frame(WHID_all = unique(deer_2018$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2018$WHID_all), by = 1))

deer_2018<-merge(deer_2018, sequence_2018, by = "occassion")
deer_2018<-merge(deer_2018, trap_names, by = "SiteID")
deer_2018<-merge(deer_2018, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2018<-deer_2018%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
        trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2018<-usage_2018%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2018<-merge(cams_2018, trap_names, by = "SiteID")
cams_2018<-cams_2018%>%
  dplyr::select(trap.id = trap_id_number, x, y)

x<-st_as_sf(cams_2018, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2018$trap.id)

#And merge the new coordinates back into the cam dataframe
cams_2018<-merge(cams_2018, coordinates, by = "trap.id")
cams_2018<-cams_2018%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2018
#Just need to collapse across all the occassions 
cams_2018$usage<-apply( usage_2018[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2018$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2018
#Need to clean it up a bit
unmarked_2018<-unmarked_2018%>%
  mutate(session = "Oakbay2018")%>%
  dplyr::select(session, 2:ncol(unmarked_2018))

#Fix the column names
colnames(unmarked_2018)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2019 CAPTURE FILES
```{r}
#2019 has dates going from October 8 to November 11th, so I am going to process this as normal and then chop it down to 30 occassions afterwards

#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2019<-data.frame(occassion = seq(ymd("2019-10-08"), ymd("2019-11-11"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:35))#These are the number we want to replace them with

unique(deer_2019$SiteID)

#Make sure occassion is read in as a date
deer_2019<-deer_2019%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

unique(deer_2019$SiteID)

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2019<- data.frame(SiteID= usage_2019$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2019$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2019<-data.frame(WHID_all = unique(deer_2019$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2019$WHID_all), by = 1))

deer_2019<-merge(deer_2019, sequence_2019, by = "occassion")
deer_2019<-merge(deer_2019, trap_names, by = "SiteID")
deer_2019<-merge(deer_2019, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2019<-deer_2019%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2019<-usage_2019%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2019<-merge(cams_2019, trap_names, by = "SiteID")
cams_2019<-cams_2019%>%
  dplyr::select(trap.id = trap_id_number, x, y)

#library(sf)
x<-st_as_sf(cams_2019, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2019$trap.id)



#And merge the new coordinates back into the cam dataframe
cams_2019<-merge(cams_2019, coordinates, by = "trap.id")
cams_2019<-cams_2019%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2019
#Just need to collapse across all the occassions 
cams_2019$usage<-apply( usage_2019[, 4:38], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2019$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2019
unmarked_2019
#Need to clean it up a bit
unmarked_2019<-unmarked_2019%>%
  mutate(session = "Oakbay2019")%>%
  dplyr::select(session, 2:ncol(unmarked_2019))

#Fix the column names
colnames(unmarked_2019)<-c("session", 1:35)

```

Modifying the files for the Multisession input
```{r}
#Ok because 2019 was a weird year and we went from October 8th to November 11th, we have 35 occassions. But to make this play nice with the other years where we have 30 occassions, I need to modify the prepped files

#remove any occassions above 30 for the capture file
deer_2019<-deer_2019%>%
  filter(occassion <= 30)

#Remove occassions from the trap file, will just have to trim the string
cams_2019<-cams_2019%>%
  mutate(usage = str_sub(usage, start = 1, end = 30))

#And then for the unmarked matrix we just remove the columns named 31-35
unmarked_2019<-unmarked_2019%>%
  dplyr::select(-c("31", "32", "33", "34", "35"))


```

#######################################################################################
#######################################################################################
#######################################################################################


FORMATTING 2020 CAPTURE FILES
```{r}
#2020 has dates going from October 8 to November 11th, so I am going to process this as normal and then chop it down to 30 occassions afterwards

#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2020<-data.frame(occassion = seq(ymd("2020-10-08"), ymd("2020-11-11"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:35))#These are the number we want to replace them with

unique(deer_2020$SiteID)

#Make sure occassion is read in as a date
deer_2020<-deer_2020%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

unique(deer_2020$SiteID)

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2020<- data.frame(SiteID= usage_2020$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2020$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2020<-data.frame(WHID_all = unique(deer_2020$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2020$WHID_all), by = 1))

deer_2020<-merge(deer_2020, sequence_2020, by = "occassion")
deer_2020<-merge(deer_2020, trap_names, by = "SiteID")
deer_2020<-merge(deer_2020, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2020<-deer_2020%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
         trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2020<-usage_2020%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2020<-merge(cams_2020, trap_names, by = "SiteID")
cams_2020<-cams_2020%>%
  dplyr::select(trap.id = trap_id_number, x, y)

#library(sf)
x<-st_as_sf(cams_2020, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2020$trap.id)



#And merge the new coordinates back into the cam dataframe
cams_2020<-merge(cams_2020, coordinates, by = "trap.id")
cams_2020<-cams_2020%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2020
#Just need to collapse across all the occassions 
cams_2020$usage<-apply( usage_2020[, 4:38], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2020$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2020
unmarked_2020
#Need to clean it up a bit
unmarked_2020<-unmarked_2020%>%
  mutate(session = "Oakbay2020")%>%
  dplyr::select(session, 2:ncol(unmarked_2020))

#Fix the column names
colnames(unmarked_2020)<-c("session", 1:35)

```

Modifying the files for the Multisession input
```{r}
#Ok because 2020 was a weird year and we went from October 8th to November 11th, we have 35 occassions. But to make this play nice with the other years where we have 30 occassions, I need to modify the prepped files

#remove any occassions above 30 for the capture file
deer_2020<-deer_2020%>%
  filter(occassion <= 30)

#Remove occassions from the trap file, will just have to trim the string
cams_2020<-cams_2020%>%
  mutate(usage = str_sub(usage, start = 1, end = 30))

#And then for the unmarked matrix we just remove the columns named 31-35
unmarked_2020<-unmarked_2020%>%
  dplyr::select(-c("31", "32", "33", "34", "35"))


```


#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2021 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2021<-data.frame(occassion = seq(ymd("2021-09-01"), ymd("2021-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2021<-deer_2021%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2021<- data.frame(SiteID= usage_2021$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2021$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2021<-data.frame(WHID_all = unique(deer_2021$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2021$WHID_all), by = 1))

deer_2021<-merge(deer_2021, sequence_2021, by = "occassion")
deer_2021<-merge(deer_2021, trap_names, by = "SiteID")
deer_2021<-merge(deer_2021, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2021<-deer_2021%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
        trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2021<-usage_2021%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2021<-merge(cams_2021, trap_names, by = "SiteID")
cams_2021<-cams_2021%>%
  dplyr::select(trap.id = trap_id_number, x, y)

x<-st_as_sf(cams_2021, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2021$trap.id)

#And merge the new coordinates back into the cam dataframe
cams_2021<-merge(cams_2021, coordinates, by = "trap.id")
cams_2021<-cams_2021%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2021
#Just need to collapse across all the occassions 
cams_2021$usage<-apply( usage_2021[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2021$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2021
#Need to clean it up a bit
unmarked_2021<-unmarked_2021%>%
  mutate(session = "Oakbay2021")%>%
  dplyr::select(session, 2:ncol(unmarked_2021))

#Fix the column names
colnames(unmarked_2021)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################



FORMATTING 2022 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2022<-data.frame(occassion = seq(ymd("2022-09-01"), ymd("2022-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2022<-deer_2022%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2021<- data.frame(SiteID= usage_2021$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2021$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2021<-data.frame(WHID_all = unique(deer_2021$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2021$WHID_all), by = 1))

deer_2022<-merge(deer_2022, sequence_2022, by = "occassion")
deer_2022<-merge(deer_2022, trap_names, by = "SiteID")
deer_2022<-merge(deer_2022, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2022<-deer_2022%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
        trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2022<-usage_2022%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2022<-merge(cams_2022, trap_names, by = "SiteID")
cams_2022<-cams_2022%>%
  dplyr::select(trap.id = trap_id_number, x, y)

x<-st_as_sf(cams_2022, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2022$trap.id)

#And merge the new coordinates back into the cam dataframe
cams_2022<-merge(cams_2022, coordinates, by = "trap.id")
cams_2022<-cams_2022%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2022
#Just need to collapse across all the occassions 
cams_2022$usage<-apply( usage_2022[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2022$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2022
#Need to clean it up a bit
unmarked_2022<-unmarked_2022%>%
  mutate(session = "Oakbay2022")%>%
  dplyr::select(session, 2:ncol(unmarked_2022))

#Fix the column names
colnames(unmarked_2022)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################

#######################################################################################
#######################################################################################
#######################################################################################

FORMATTING 2023 CAPTURE FILES
```{r}
#First we need to fix the occassions so that they are just 1 to 30 (30 days has september)
sequence_2023<-data.frame(occassion = seq(ymd("2023-09-01"), ymd("2023-09-30"), by = "day"), #These are the days included in the file
                     occassion_number = seq(1:30))#These are the number we want to replace them with

#Make sure occassion is read in as a date
deer_2023<-deer_2023%>%
  rename(occassion = date)%>%
  mutate(occassion = ymd(occassion))

#create a frame of trap names from the usage file (remember detections might not have all the traps!)
#trap_names_2023<- data.frame(SiteID= usage_2021$SiteID,
 #                       trap.id_number = seq(from = 1, to = n_distinct(usage_2021$SiteID), by = 1))

#Create a frame of wildlife ID numbers
#wildlife_ids_2021<-data.frame(WHID_all = unique(deer_2021$WHID_all),
 #                        id_number = seq(from = 1, to = n_distinct(deer_2021$WHID_all), by = 1))

deer_2023<-merge(deer_2023, sequence_2023, by = "occassion")
deer_2023<-merge(deer_2023, trap_names, by = "SiteID")
deer_2023<-merge(deer_2023, id_names, by = "WHID_all")

#And then finally subset the numbered columns we want, and order them properly
deer_2023<-deer_2023%>%
  dplyr::select(session, 
         id = id_number, 
         occassion = occassion_number, 
        trap.id = trap_id_number)
```

Formatting the trap files
```{r}
#Clean up the location file
cams_2023<-usage_2023%>%
  dplyr::select(c(SiteID, 
           x = easting, y = northing))
cams_2023<-merge(cams_2023, trap_names, by = "SiteID")
cams_2023<-cams_2023%>%
  dplyr::select(trap.id = trap_id_number, x, y)

x<-st_as_sf(cams_2023, coords = c("y", "x"), crs = 4326)
# Transform the coordinates to UTM zone 10
x_utm <- st_transform(x, crs = "+proj=utm +zone=10 +datum=WGS84 +units=m")
# Extract the coordinates and add trap IDs
coordinates <- st_coordinates(x_utm)
coordinates <- as.data.frame(coordinates)
coordinates <- coordinates %>%
  mutate(trap.id = cams_2023$trap.id)

#And merge the new coordinates back into the cam dataframe
cams_2023<-merge(cams_2023, coordinates, by = "trap.id")
cams_2023<-cams_2023%>%
  dplyr::select(-c(x, y))%>%
  dplyr::select(trap.id, x = X, y= Y) #I dont know WHY, but the x and y come out backwards...

#Now need a usage column - this is a series of 1's and 0's which tells secr whether
#the camera was on each day. We can use the usage matrix for this
usage_2023
#Just need to collapse across all the occassions 
cams_2023$usage<-apply( usage_2023[, 4:33], 1, paste, collapse = "")
#Check to make sure that everyone is the same length
str_length(cams_2023$usage)


```

Unmarked matrix
```{r}
#TU - unmarked matrix 2018
unmarked_2023
#Need to clean it up a bit
unmarked_2023<-unmarked_2023%>%
  mutate(session = "Oakbay2023")%>%
  dplyr::select(session, 2:ncol(unmarked_2023))

#Fix the column names
colnames(unmarked_2023)<-c("session", 1:30)

```
#######################################################################################
#######################################################################################
#######################################################################################


And now combining the files so that they can be exported for a multisession model
```{r}

#For the capture files, we just create a single file
deer_multi<-rbind(deer_2018, deer_2019, deer_2020, deer_2021, deer_2022, deer_2023)

write.table(deer_multi, file.path(output_directory, "deer_multi.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Same with the unmarked matrices
unmarked_multi<-rbind(unmarked_2018, unmarked_2019, unmarked_2020, unmarked_2021, unmarked_2022, unmarked_2023)

write.table(unmarked_multi, file.path(output_directory, "unmarked_multi.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#And then the trap files are exported as normal
write.table(cams_2018, file.path(output_directory, "deer_cameras_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(cams_2019, file.path(output_directory, "deer_cameras_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(cams_2020, file.path(output_directory, "deer_cameras_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(cams_2021, file.path(output_directory, "deer_cameras_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(cams_2022, file.path(output_directory, "deer_cameras_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(cams_2023, file.path(output_directory, "deer_cameras_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

#Write the availability files
write.table(markocc_2018, file.path(output_directory, "markocc_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(markocc_2019, file.path(output_directory, "markocc_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(markocc_2020, file.path(output_directory, "markocc_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(markocc_2021, file.path(output_directory, "markocc_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(markocc_2022, file.path(output_directory, "markocc_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(markocc_2023, file.path(output_directory, "markocc_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


```

#####################################################################################################
Extra bit here for outputting the files for individual year analysis
```{r}
#Ok first, we can write the main detection file for 2018
write.table(deer_2018, file.path(output_directory, "deer_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2018, file.path(output_directory, "unmarked_2018.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_2019, file.path(output_directory, "deer_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2019, file.path(output_directory, "unmarked_2019.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_2020, file.path(output_directory, "deer_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2020, file.path(output_directory, "unmarked_2020.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_2021, file.path(output_directory, "deer_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2021, file.path(output_directory, "unmarked_2021.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_2022, file.path(output_directory, "deer_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2022, file.path(output_directory, "unmarked_2022.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(deer_2023, file.path(output_directory, "deer_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

write.table(unmarked_2023, file.path(output_directory, "unmarked_2023.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#Next, lets find a way to add the unsighted individuals. So we have this dataframe, which tells us whether all the individuals in the study were available for deteciton or not on our specified dates. 
markocc_2018

#So in this dataframe, we only want to keep individuals that were available, 
markocc_2018<-markocc_2018 %>%
  filter(if_any(2:31, ~ . == 1))

#Ok now match the names
known<-merge(markocc_2018, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))

#Ok, NOW, which of these deer were NOT seen in the capture file? 
known<-known%>%
  filter(!id %in% unique(deer_2018$id))

#So now we append them to the deer_2019 file? And give them dummy variables?
known<-known%>%
  select(id)%>%
  mutate(session = "Oakbay2018",
         occassion = 0,
         trap.id = 1) #Arbitrary trap location

deer_2018_known<-rbind(deer_2018, known)

#Write this and lets try it
write.table(deer_2018_known, file.path(output_directory, "deer_2018_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs


#Ok lets do it for the rest of the years now
markocc_2019<-markocc_2019%>%
  filter(if_any(2:31, ~ . == 1))

known<-merge(markocc_2019, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))%>%
  filter(!id %in% unique(deer_2019$id))%>%
  select(id)%>%
  mutate(session = "Oakbay2019",
         occassion = 0,
         trap.id = 2) #Arbitrary trap location

deer_2019_known<-rbind(deer_2019, known)
################################################
markocc_2020<-markocc_2020%>%
  filter(if_any(2:31, ~ . == 1))

known<-merge(markocc_2020, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))%>%
  filter(!id %in% unique(deer_2020$id))%>%
  select(id)%>%
  mutate(session = "Oakbay2020",
         occassion = 0,
         trap.id = 1) #Arbitrary trap location

deer_2020_known<-rbind(deer_2020, known)
################################################
markocc_2021<-markocc_2021%>%
  filter(if_any(2:31, ~ . == 1))

known<-merge(markocc_2021, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))%>%
  filter(!id %in% unique(deer_2021$id))%>%
  select(id)%>%
  mutate(session = "Oakbay2021",
         occassion = 0,
         trap.id = 1) #Arbitrary trap location

deer_2021_known<-rbind(deer_2021, known)
################################################
markocc_2022<-markocc_2022%>%
  filter(if_any(2:31, ~ . == 1))

known<-merge(markocc_2022, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))%>%
  filter(!id %in% unique(deer_2022$id))%>%
  select(id)%>%
  mutate(session = "Oakbay2022",
         occassion = 0,
         trap.id = 1) #Arbitrary trap location

deer_2022_known<-rbind(deer_2022, known)
################################################
markocc_2023<-markocc_2023%>%
  filter(if_any(2:31, ~ . == 1))

known<-merge(markocc_2023, id_names, by = c("WHID_all"))%>%
  relocate(id_number)%>%
  rename(id = id_number)%>%
  select(-c(WHID_all))%>%
  filter(!id %in% unique(deer_2023$id))%>%
  select(id)%>%
  mutate(session = "Oakbay2023",
         occassion = 0,
         trap.id = 1) #Arbitrary trap location

deer_2023_known<-rbind(deer_2023, known)

write.table(deer_2019_known, file.path(output_directory, "deer_2019_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(deer_2020_known, file.path(output_directory, "deer_2020_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(deer_2021_known, file.path(output_directory, "deer_2021_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(deer_2022_known, file.path(output_directory, "deer_2022_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs
write.table(deer_2023_known, file.path(output_directory, "deer_2023_known.txt"), sep = " ", 
            quote = FALSE,     #no quotes allowed!
            row.names = FALSE) #prevent including row IDs

```